\documentclass[12pt,a4paper]{article}

% ---- Packages ----
\usepackage[utf8]{inputenc}
\usepackage[T1]{fontenc}
\usepackage{amsmath,amssymb,amsthm}
\usepackage{mathtools}
\usepackage{physics}
\usepackage{geometry}
\usepackage{hyperref}
\usepackage{cleveref}
\usepackage{graphicx}
\usepackage{booktabs}
\usepackage{array}
\usepackage{longtable}
\usepackage{enumitem}
\usepackage{listings}
\usepackage{xcolor}
\usepackage{tikz-cd}

\geometry{margin=1in}
\hypersetup{
    colorlinks=true,
    linkcolor=blue!70!black,
    citecolor=green!50!black,
    urlcolor=blue!60!black
}

% ---- Listings styles ----
\lstdefinestyle{python}{
    language=Python,
    basicstyle=\ttfamily\small,
    keywordstyle=\color{blue!70!black}\bfseries,
    commentstyle=\color{green!50!black}\itshape,
    stringstyle=\color{red!60!black},
    showstringspaces=false,
    breaklines=true,
    frame=single,
    backgroundcolor=\color{gray!5},
    numbers=left,
    numberstyle=\tiny\color{gray},
    xleftmargin=2em,
    framexleftmargin=1.5em
}

\lstdefinestyle{scala}{
    language=Scala,
    basicstyle=\ttfamily\small,
    keywordstyle=\color{blue!70!black}\bfseries,
    commentstyle=\color{green!50!black}\itshape,
    stringstyle=\color{red!60!black},
    morekeywords={val,def,case,class,object,trait,sealed,extends,given,override,open,new,with,import,match,type},
    showstringspaces=false,
    breaklines=true,
    frame=single,
    backgroundcolor=\color{gray!5},
    numbers=left,
    numberstyle=\tiny\color{gray},
    xleftmargin=2em,
    framexleftmargin=1.5em,
    literate={=>}{{\textcolor{blue!70!black}{=>}}}2
             {>>>}{{\textcolor{red!70!black}{\texttt{>{>}>}}}}3
             {&&&}{{\textcolor{red!70!black}{\texttt{\&\&\&}}}}3
}

% ---- Theorem environments ----
\theoremstyle{definition}
\newtheorem{definition}{Definition}[section]
\newtheorem{theorem}{Theorem}[section]
\newtheorem{proposition}{Proposition}[section]
\newtheorem{remark}{Remark}[section]

% ---- Title ----
\title{\textbf{Topological Hydrodynamics} \\[0.5em]
\large From Proton Spin Asymmetry to Programmable Spacetime}
\author{Wyatt Meldman-Floch}
\date{\copyright\ 2026}

\begin{document}
\maketitle

\begin{abstract}
We present a unified theoretical framework connecting the proton spin crisis,
the Sivers asymmetry, spin network renormalization, and topological quantum
computation. Beginning with the experimental observation that quark intrinsic
spins account for only $\approx 30\%$ of the proton's total spin $J = 1/2$,
we develop a categorical and cohomological interpretation in which the
``missing'' angular momentum is encoded in the topology of an underlying
spin network. The Sivers function is identified as a co-boundary obstruction
in a Poincar\'{e} complex, the renormalization group flow is cast as a
functorial coarse-graining via the Day convolution, and the proton itself
is modeled as a self-correcting homological code. We propose a quantum
circuit architecture---built on Majorana zero modes and surface codes---that
programs torsion directly into the lattice geometry, enabling the simulation
of traversable wormholes through syndrome teleportation. A distributed rApp
protocol---implemented in the Babel categorical DSL on the Reality
SDK---is formulated to govern topological charge conservation across
entangled Poincar\'{e} complexes, and the critical exponents for the
traversability phase transition are estimated against modern quantum hardware
capabilities.
\end{abstract}

\tableofcontents
\newpage

%==========================================================================
\section{The Proton Spin Crisis}\label{sec:spin-crisis}
%==========================================================================

\subsection{Experimental Origins}

Proton spin asymmetry measures how the scattering of protons depends on
their spin orientation relative to particle beams. The 1987 European Muon
Collaboration (EMC) experiment revealed that quark intrinsic spins contribute
only a small fraction---roughly $12$--$30\%$---of the total proton spin,
contradicting the na\"{\i}ve quark-model expectation that three valence
quarks carry the bulk of the angular momentum. This discrepancy became known
as the \emph{proton spin crisis}.

The total proton spin decomposes as
\begin{equation}\label{eq:spin-sum}
    J = \frac{1}{2}
    = \frac{1}{2}\Delta\Sigma + \Delta G + L_q + L_g \,,
\end{equation}
where $\Delta\Sigma$ is the quark spin contribution, $\Delta G$ the gluon
spin contribution, and $L_q$, $L_g$ the orbital angular momentum of quarks
and gluons respectively.

\subsection{Key Observables}

\begin{description}[leftmargin=1.5cm,style=nextline]
    \item[Longitudinal Asymmetry $A_1$]
    Measured in deep-inelastic scattering (DIS) of polarized leptons off
    polarized protons, this observable reveals the helicity distributions
    of quarks and gluons. The integral of the spin-dependent structure
    function $g_1$ constrains $\Delta\Sigma$.

    \item[Transverse Single-Spin Asymmetry (TSSA)]
    Observed in proton--proton collisions, these asymmetries are often
    unexpectedly large, indicating non-zero orbital angular momentum
    and the need for transverse-momentum-dependent (TMD) parton
    distribution functions.
\end{description}

\subsection{Experimental Facilities}

Research at Brookhaven National Laboratory's Relativistic Heavy Ion Collider
(RHIC) has been instrumental in constraining the gluon spin contribution
$\Delta G$. The future Electron-Ion Collider (EIC) will map the
three-dimensional structure of the proton with unprecedented precision,
probing generalized parton distributions (GPDs) and TMDs simultaneously.

%==========================================================================
\section{Spin Network Interpretation}\label{sec:spin-networks}
%==========================================================================

\subsection{The Proton as a Bound State of a Spin Network}

In Loop Quantum Gravity (LQG), space is discretized into a \emph{spin
network}---a graph whose edges carry spin representations
$j = 1/2, 1, 3/2, \ldots$ and whose nodes represent quanta of volume.

\begin{definition}[Spin-Network Proton]
Rather than viewing a proton as a point-like particle propagating through a
background spacetime, we model it as a topological defect---a specific
excitation of the spin network. The total spin $J = 1/2$ is an emergent
property of the local geometry and the way the network edges are woven
together.
\end{definition}

\subsection{The Crisis as a Geometric Phenomenon}

From the spin-network perspective, the ``missing'' spin admits a natural
explanation:
\begin{itemize}
    \item \textbf{Relational entanglement.} In a spin network, angular
    momentum is a property of the connections (edges) between nodes. If
    quarks are nodes, much of the angular momentum is stored in the edges
    (gluon fields / geometry) connecting them.

    \item \textbf{Orbital angular momentum.} Spin networks naturally
    incorporate the spatial relationships between components. What QCD calls
    ``orbital angular momentum'' is the quantum of area and volume encoded
    in the network's geometry.
\end{itemize}

\subsection{Theoretical Frameworks}

\begin{table}[ht]
\centering
\renewcommand{\arraystretch}{1.3}
\begin{tabular}{@{}p{4cm}p{9cm}@{}}
\toprule
\textbf{Framework} & \textbf{Perspective on Proton Spin} \\
\midrule
Braided Matter Models &
Elementary particles are ``braids'' in the spin network
(Bilson-Thompson). Spin is a topological property of how the
strands twist. \\
Holographic Spin Networks &
Internal degrees of freedom are holographically encoded on a 2D
surface. Asymmetry arises from the boundary map. \\
Algebraic QFT &
Uses spin-foam logic to describe the flux of spin, naturally
accommodating the gluon contribution. \\
\bottomrule
\end{tabular}
\caption{Theoretical frameworks connecting spin networks to proton structure.}
\label{tab:frameworks}
\end{table}

%==========================================================================
\section{The Sivers Effect and Functorial Models}\label{sec:sivers}
%==========================================================================

\subsection{The Sivers Function}

The Sivers function $f_{1T}^{\perp q}(x, \vb{k}_\perp^2)$ describes
the correlation between the proton's transverse spin $\vb{S}_T$ and the
quark's transverse momentum $\vb{k}_\perp$. It is \emph{T-odd}: it
changes sign between semi-inclusive DIS (SIDIS) and Drell--Yan (DY)
processes, a prediction confirmed by the COMPASS experiment.

\subsection{Gauge Links as Functors}

The Sivers effect relies on initial-/final-state interactions (ISI/FSI),
represented in QCD by Wilson lines---path-ordered exponentials of the
gauge field.

\begin{definition}[Wilson-Line Functor]
A Wilson line defines a functor from the \emph{path groupoid}
(the paths a quark traverses) to the \emph{category of vector spaces}
(the color/spin states of the quark). The Sivers asymmetry measures
the failure of this functor to be ``flat'' under a change in spin
orientation: it maps a change in the spin category to a displacement in
the momentum category.
\end{definition}

\subsection{Braided Categories and the Sign Change}

In braided-ribbon-network models, the Sivers effect is represented by a
\emph{braiding morphism}. Treating the proton spin as a topological
winding number, the asymmetry arises naturally from the braid twisting
as a parton is ejected. This picture is qualitatively consistent with the
observed sign change between SIDIS and DY.

\subsection{AdS/QCD and Light-Front Holography}

The most data-consistent ``functor-like'' mapping comes from AdS/CFT.
Light-Front Holography (LFH), developed by Brodsky and de~T\'{e}ramond,
creates a dictionary (functor) between a 5D gravitational theory and 4D
QCD. Predictions for the Sivers-like distribution of quarks align with
experimental extractions from HERMES and COMPASS.

%==========================================================================
\section{Light-Front Holographic Equations of Motion}\label{sec:lfh}
%==========================================================================

\subsection{The Light-Front Schr\"{o}dinger Equation}

The radial equation of motion derived from the AdS/QCD mapping determines
the mass spectrum of the proton:
\begin{equation}\label{eq:lf-schrodinger}
    \left(
        -\frac{d^2}{d\zeta^2}
        - \frac{1 - 4L^2}{4\zeta^2}
        + U(\zeta)
    \right) \phi(\zeta) = M^2 \phi(\zeta) \,,
\end{equation}
where $\zeta = \sqrt{x(1-x)}\,|\vb{b}_\perp|$ is the holographic
variable, $L$ is the orbital angular momentum, and the confinement
potential is
\begin{equation}
    U(\zeta) = \kappa^4 \zeta^2 + 2\kappa^2(L + S - 1) \,.
\end{equation}

\subsection{Universal Radial Profile}

The solution for a state with quantum numbers $(n, L)$ is
\begin{equation}\label{eq:radial-profile}
    \varphi_{n,L}(\zeta)
    = \kappa^{1+L}
      \sqrt{\frac{2\,n!}{(n+L)!}}\;
      \zeta^{1/2+L}\,
      e^{-\kappa^2 \zeta^2 / 2}\,
      L_n^{L}\!\bigl(\kappa^2 \zeta^2\bigr) \,.
\end{equation}
For the ground-state proton ($n=0$, $L=0$), this reduces to a
Gaussian-like profile.

\subsection{Spin-Improved Light-Front Wavefunctions}

The Sivers asymmetry requires interference between $L=0$ and $L=1$ states.
The spin-improved wavefunctions for a quark $q$ inside a proton are
\begin{align}
    \psi_{+,\,h=+1/2}^{q}(x, \vb{k}_\perp)
    &= \varphi_{L=0}(x, \vb{k}_\perp)
       + \alpha^q \frac{\vb{k}_\perp^2}{M^2}\,
         \varphi_{L=1}(x, \vb{k}_\perp) \,,
    \label{eq:lfwf-plus} \\[6pt]
    \psi_{-,\,h=+1/2}^{q}(x, \vb{k}_\perp)
    &= -\frac{k^1 - ik^2}{xM}\,
       \beta^q\,\varphi_{L=1}(x, \vb{k}_\perp) \,,
    \label{eq:lfwf-minus}
\end{align}
where $\alpha^q$ and $\beta^q$ are parameters fixed by the anomalous
magnetic moments of the proton and neutron.

\subsection{Extraction of the Sivers Function}

The Sivers function is computed as the overlap integral
\begin{equation}\label{eq:sivers-extraction}
    f_{1T}^{\perp q}(x, \vb{k}_\perp^2)
    = \frac{M}{2\pi}\,
      \operatorname{Im}\!\left[
          \psi_{+,+1/2}^{q*}(x, \vb{k}_\perp)\;
          \psi_{-,+1/2}^{q}(x, \vb{k}_\perp)
      \right] ,
\end{equation}
yielding the asymmetric ``dipole'' distribution observed in COMPASS and
HERMES data: positive for $u$-quarks, negative for $d$-quarks.

%==========================================================================
\section{Unifying Spin-Foam Renormalization with Light-Front Evolution}
\label{sec:unification}
%==========================================================================

\subsection{Scale as Resolution}

In light-front evolution, the renormalization scale $Q^2$ sets the
resolution of the proton's constituents. In spin foams, scale is replaced
by the refinement limit of the discretization. The unifying insight is:

\begin{proposition}[Functorial Link]
The boosted light-front time $x^+$ acts as a foliation parameter for the
spin foam. As one evolves along the light front, the spin foam undergoes
Pachner moves that increase the number of edges and nodes (quarks and gluons).
\end{proposition}

\subsection{RG Flow as Coarse-Graining}

\begin{itemize}
    \item \textbf{Light-front view.} The RG flow (via DGLAP or CSS
    equations) evolves the TMD distributions.

    \item \textbf{Spin-foam view.} The RG flow is a coarse-graining
    procedure on the spin network. Tensor Network Renormalization (TNR,
    see \cref{sec:cg-operations} for the detailed formalism)
    shows how a complex web of spins (high $Q^2$) averages into a
    single effective ``valence quark'' node at low energy.

    \item \textbf{Synthesis.} The equations of motion for the Sivers
    effect emerge when \emph{cylindrical consistency} of the spin network
    (required for diffeomorphism invariance; see
    \cref{sec:cylindrical}) is forced to match the
    \emph{unitarity constraints} of the light-front Hamiltonian.
\end{itemize}

\subsection{Hamiltonian Constraint in Light-Front Coordinates}

The Hamiltonian constraint from LQG, $\hat{H}\Psi = 0$, is projected
onto the light-front eigenvalue problem:
\begin{equation}\label{eq:lf-eigenvalue}
    \hat{H}_{LF}\ket{\Psi_{\text{proton}}} = M^2 \ket{\Psi_{\text{proton}}} \,.
\end{equation}
In a unified model, $\hat{H}_{LF}$ is derived from the vertex amplitudes
of the spin foam. The Sivers asymmetry appears as a non-commutative phase
in the vertex amplitude when the spin labels on the network edges are
polarized.

%==========================================================================
\section{Spin Network Coarse-Graining and the Proton as a Complex Network}
\label{sec:coarse-graining}
%==========================================================================

The preceding sections invoke coarse-graining of spin networks
(\cref{sec:unification}) and iterated hashing
(\cref{sec:latency}) informally. This section provides the
mathematical backbone: we define the colored spin network of the
proton, specify three coarse-graining operations, state the
cylindrical-consistency condition that connects scales, and show how
diffeomorphism invariance is restored in the continuum limit.
The proton then appears as a multi-scale complex network whose
topological complexity is directly related to its mass.

%------------------------------------------------------------------
\subsection{The Colored Spin Network of the Proton}
\label{sec:colored-spin-network}
%------------------------------------------------------------------

Let $\Gamma = (V, E)$ be a finite oriented graph embedded in
a spatial 3-manifold $\Sigma$. To each edge $e \in E$ we assign
an irreducible representation $\rho_e$ of $\mathrm{SU}(3)_c$
(the color gauge group) and, independently, a spin
$j_e \in \{0, 1/2, 1, \ldots\}$ of $\mathrm{SU}(2)_J$
(rotational symmetry). At every node $v \in V$ we place an
intertwiner
\begin{equation}
    \iota_v \;\in\;
    \mathrm{Inv}\!\Bigl(
        \bigotimes_{e \ni v} \rho_e \otimes V_{j_e}
    \Bigr) ,
\end{equation}
where the tensor product runs over all edges incident at $v$ and
$\mathrm{Inv}(\cdot)$ denotes the invariant subspace under the
joint action of $\mathrm{SU}(3)_c \times \mathrm{SU}(2)_J$.

\begin{definition}[Colored Spin-Network Hilbert Space]
\label{def:colored-hilbert}
The Hilbert space of colored spin-network states on $\Gamma$ is
\begin{equation}\label{eq:colored-hilbert}
    \mathcal{H}_\Gamma
    \;=\;
    \bigoplus_{\{\rho_e, j_e\}}
    \;\bigotimes_{v \in V} \mathcal{I}_v
    \;\otimes\;
    \bigotimes_{e \in E} V_{\rho_e} \otimes V_{j_e} \,,
\end{equation}
where $\mathcal{I}_v$ is the intertwiner space at node $v$
and the direct sum runs over all admissible labelings.
\end{definition}

\begin{remark}[Connection to the chain complex]
The nodes $V$ and edges $E$ of $\Gamma$ are precisely the
$0$-cells $C_0$ and $1$-cells $C_1$ of the chain complex
introduced in \cref{sec:boundary}: quarks sit at $C_0$ and
gluon flux tubes span $C_1$. The intertwiner $\iota_v$ at each
node enforces the local color-singlet constraint
(cf.\ the closure condition of
\cref{def:poincare-protocol}).
\end{remark}

%------------------------------------------------------------------
\subsection{Coarse-Graining Operations}
\label{sec:cg-operations}
%------------------------------------------------------------------

Three elementary operations reduce the complexity of
$(\Gamma, \{\rho_e, j_e, \iota_v\})$ while preserving
long-range physics.

\paragraph{1. Edge contraction.}
Given an edge $e = (v_1, v_2)$, contract it to a single node
$v_{12}$ with effective intertwiner
\begin{equation}\label{eq:edge-contraction}
    \iota_{v_{12}}^{\mathrm{eff}}
    \;=\;
    \sum_{\rho_e, j_e}
    d_{\rho_e}\, d_{j_e}\;
    \bigl(\iota_{v_1} \otimes \iota_{v_2}\bigr)
    \big|_{\mathrm{Inv}} \,,
\end{equation}
where $d_{\rho}$ and $d_j$ are the dimensions of the
respective representations and the projection is onto the
invariant subspace. This merges two partons into one effective
degree of freedom.

\paragraph{2. Node decimation (block spin).}
Partition $V$ into disjoint blocks $\{B_\alpha\}$. For each
block, trace over all internal edges:
\begin{equation}\label{eq:node-decimation}
    \mathcal{I}_{B_\alpha}^{\mathrm{eff}}
    \;=\;
    \operatorname{Tr}_{E_{\mathrm{int}}(B_\alpha)}
    \!\Bigl(
        \bigotimes_{v \in B_\alpha} \iota_v
    \Bigr) .
\end{equation}
The result is a coarse graph $\Gamma'$ with one supernode per
block. In QCD language this replaces a sea-quark--gluon
subcloud by a single constituent quark.

\paragraph{3. Tensor network renormalization (TNR).}
Write the partition function as a tensor
contraction~\cite{EvenblyVidal2015, DittrichMizera2016}:
\begin{equation}\label{eq:tnr}
    Z_\Gamma
    \;=\;
    \sum_{\{j_e, \rho_e\}}
    \prod_{v \in V} T_v^{(\iota_v)}
    \;\prod_{e \in E} d_{\rho_e}\, d_{j_e} \,,
\end{equation}
where $T_v^{(\iota_v)}$ is the intertwiner tensor at node $v$.
Perform a singular-value decomposition on every pair of adjacent
tensors, truncate to the leading $\chi$ singular values, and
reconstitute the network on a coarser graph~$\Gamma'$.

\begin{table}[ht]
\centering
\renewcommand{\arraystretch}{1.3}
\begin{tabular}{@{}p{3.2cm}p{4.0cm}p{5.8cm}@{}}
\toprule
\textbf{CG Operation} &
\textbf{Spin-Network Action} &
\textbf{QCD / Proton Interpretation} \\
\midrule
Edge contraction &
Merge two nodes via an edge; sum over internal reps &
Quark--gluon fusion: two partons merge into one
effective parton \\
Node decimation &
Trace over internal edges of a block &
Constituent-quark picture: a sea-quark cloud collapses
into a single constituent quark \\
TNR (SVD truncation) &
Contract and truncate tensor network &
DGLAP / CSS evolution: integrate out high-$k_\perp$
modes above a cutoff $\chi$ \\
\bottomrule
\end{tabular}
\caption{Coarse-graining operations and their QCD counterparts.}
\label{tab:cg-operations}
\end{table}

%------------------------------------------------------------------
\subsection{Cylindrical Consistency and the Embedding Maps}
\label{sec:cylindrical}
%------------------------------------------------------------------

Let $\Gamma \subset \Gamma'$ denote a coarsening
($\Gamma$ is obtained from $\Gamma'$ by one or more of the
operations above). Define the \emph{embedding map}
$p_{\Gamma'\!\to\Gamma}
  : \mathcal{H}_{\Gamma'} \to \mathcal{H}_\Gamma$
by partial trace over the degrees of freedom in
$\Gamma' \setminus \Gamma$.

\begin{definition}[Cylindrical Consistency]
\label{def:cylindrical-consistency}
A family of states
$\{\Psi_\Gamma \in \mathcal{H}_\Gamma\}_\Gamma$ is
\emph{cylindrically consistent} if, for every pair
$\Gamma \subset \Gamma'$,
\begin{equation}\label{eq:cylindrical-consistency}
    p_{\Gamma'\!\to\Gamma}\,\Psi_{\Gamma'}
    \;=\; \Psi_\Gamma \,.
\end{equation}
\end{definition}

The condition states that coarse-graining the fine state
$\Psi_{\Gamma'}$ must reproduce the coarse state
$\Psi_\Gamma$ exactly. This is the spin-network analog of
the statement that a valence-quark observable at low $Q^2$
must be reproducible from the full parton-level description
at high $Q^2$.

The consistency condition is summarized by the commutative
diagram
\begin{equation}\label{eq:cyl-diagram}
\begin{tikzcd}
    \mathcal{H}_{\Gamma''}
        \arrow[r, "p_{\Gamma''\!\to\Gamma'}"]
        \arrow[dr, swap, "p_{\Gamma''\!\to\Gamma}"]
    & \mathcal{H}_{\Gamma'}
        \arrow[d, "p_{\Gamma'\!\to\Gamma}"]
    \\
    & \mathcal{H}_{\Gamma}
\end{tikzcd}
\end{equation}
which must commute for any triple
$\Gamma \subset \Gamma' \subset \Gamma''$.

%------------------------------------------------------------------
\subsection{Restoration of Diffeomorphism Invariance}
\label{sec:diffeo}
%------------------------------------------------------------------

A spin network on a fixed graph $\Gamma$ breaks the
diffeomorphism invariance of the continuum theory. Three
complementary mechanisms restore it in the
limit~\cite{Rovelli2004, ThiemannBook2007}.

\paragraph{1. Diffeomorphism averaging.}
The physical (diffeomorphism-invariant) Hilbert space is
obtained as a quotient~\cite{ThiemannBook2007}:
\begin{equation}\label{eq:diffeo-averaging}
    \mathcal{H}_{\mathrm{phys}}
    \;=\;
    \mathcal{H}_{\mathrm{kin}} \big/ \mathrm{Diff}(\Sigma) \,,
\end{equation}
where $\mathrm{Diff}(\Sigma)$ acts by moving graph vertices
and edges through $\Sigma$. Two spin-network states that
differ only by a diffeomorphism are identified; only the
combinatorial (abstract-graph) data survive.

\paragraph{2. Perfect discretization.}
Following Bahr and Dittrich~\cite{BahrDittrich2009}, one seeks
an effective action $S_{\mathrm{eff}}[\Gamma]$ on the coarse
graph such that
\begin{equation}\label{eq:perfect-discretization}
    Z_{\Gamma}^{\mathrm{eff}}
    \;=\;
    \int \mathcal{D}\phi_{\mathrm{fine}}\;
    e^{-S[\Gamma', \phi_{\mathrm{fine}}]}
    \;=\;
    e^{-S_{\mathrm{eff}}[\Gamma]} \,,
\end{equation}
i.e.\ the coarse partition function captures the full
fine-grained dynamics exactly. The resulting discrete
theory is automatically diffeomorphism-invariant to the
extent that the continuum theory is.

\paragraph{3. Continuum limit as RG fixed point.}
Under iterated coarse-graining, the effective couplings
flow toward a fixed point~\cite{DittrichGeiller2015}:
\begin{equation}\label{eq:rg-fixedpoint}
    \lim_{n \to \infty}
    \bigl(\mathcal{R}^n \cdot S\bigr)
    \;=\;
    S^* \,,
\end{equation}
where $\mathcal{R}$ is the coarse-graining superoperator and
$S^*$ is the fixed-point action. If $S^*$ is
diffeomorphism-invariant (as expected for QCD), then every
finite truncation inherits approximate invariance, with
corrections that vanish as the lattice is refined.

\begin{remark}
For the proton, ``diffeomorphism invariance'' maps concretely
to $\mathrm{SU}(3)_c$ gauge invariance: the Wilson line
(\cref{sec:wilson-hash}) is the gauge-invariant hash that
compresses a continuous gauge-field configuration into a
single group element. The perfect-discretization condition
\eqref{eq:perfect-discretization} demands that this hash is
\emph{exact}---no gauge-invariant information is lost in the
coarse-graining.
\end{remark}

%------------------------------------------------------------------
\subsection{The Proton as a Multi-Scale Complex Network}
\label{sec:multiscale-proton}
%------------------------------------------------------------------

At different resolutions, the proton's spin network presents
qualitatively different graph topologies. \Cref{tab:multiscale-proton}
summarizes three characteristic regimes.

\begin{table}[ht]
\centering
\renewcommand{\arraystretch}{1.3}
\begin{tabular}{@{}p{2.8cm}p{3.0cm}p{3.0cm}p{4.2cm}@{}}
\toprule
\textbf{Scale} &
\textbf{Resolution ($Q^2$)} &
\textbf{Graph Topology} &
\textbf{Physical Content} \\
\midrule
UV &
$\gtrsim 100\;\mathrm{GeV}^2$ &
Dense lattice, $|V| \gg 1$ &
Sea quarks, gluons, and their OAM
(full partonic content) \\
Intermediate &
$1$--$10\;\mathrm{GeV}^2$ &
Effective graph, $|V| \sim 10$ &
Constituent quarks, gluon clusters,
valence + sea \\
IR &
$< 1\;\mathrm{GeV}^2$ &
$Y$-graph, $|V| = 3$ &
Three valence quarks connected by
flux-tube junctions \\
\bottomrule
\end{tabular}
\caption{The proton as a multi-scale complex network.}
\label{tab:multiscale-proton}
\end{table}

To quantify the network structure at each scale, introduce three
standard graph-theoretic observables:
\begin{itemize}
    \item \textbf{Degree distribution} $P(k)$: the fraction of
    nodes with $k$ incident edges.
    \item \textbf{Clustering coefficient} $C$: the probability
    that two neighbors of a node are themselves connected.
    \item \textbf{Betweenness centrality} $B_v$: the fraction
    of shortest paths in the network that pass through
    node~$v$.
\end{itemize}

\begin{proposition}[Coarse-graining simplifies network metrics]
\label{prop:network-metrics}
Under successive coarse-graining steps
$\Gamma_{\mathrm{UV}} \to \cdots \to \Gamma_{\mathrm{IR}}$:
\begin{enumerate}
    \item $P(k) \;\to\; \delta_{k,2}$\,: every node in the
    IR $Y$-graph has exactly two incident edges.
    \item $C \;\to\; 0$\,: the $Y$-graph is a tree and
    contains no triangles.
    \item $\max_v B_v \;\to\; 1/3$\,: each of the three
    valence-quark nodes lies on exactly one-third of the
    shortest paths.
\end{enumerate}
\end{proposition}

These limiting values match the topological properties of
the confining string junction that connects three quarks in
the IR limit of lattice QCD.

%------------------------------------------------------------------
\subsection{Network Complexity and Mass Generation}
\label{sec:network-mass}
%------------------------------------------------------------------

Define the \emph{topological complexity} of a colored spin
network as
\begin{equation}\label{eq:network-complexity}
    \mathcal{C}(\Gamma)
    \;=\;
    \sum_{v \in V} \ln \dim \mathcal{I}_v
    \;+\;
    \sum_{e \in E} \ln\!\bigl(d_{\rho_e}\, d_{j_e}\bigr)
    \;+\;
    \sum_{k \geq 1} \beta_k(\Gamma)\,\ln k \,,
\end{equation}
where $\beta_k(\Gamma)$ is the $k$-th Betti number of the
graph. The first term counts intertwiner degrees of freedom
(nodes / quarks), the second counts representation degrees of
freedom (edges / gluons), and the third measures topological
loops.

\begin{proposition}[Mass from complexity]
\label{prop:mass-complexity}
At the RG fixed point $\Gamma^*$ of the coarse-graining flow,
the proton mass satisfies
\begin{equation}\label{eq:mass-complexity}
    M_p^2
    \;\propto\;
    \mathcal{C}(\Gamma^*) \,,
\end{equation}
i.e.\ the squared mass is proportional to the topological
complexity of the fixed-point network. This gives a
microscopic derivation of the ``mass as complexity''
identification in the topological Einstein equation
\eqref{eq:topological-einstein} of \cref{sec:quantum-gravity}.
\end{proposition}

The connection is natural: a more complex spin network requires
a larger error-correction overhead (syndrome entropy) to
stabilize, and the stabilizer pressure is precisely the mass
seen by an external observer. The coarse-graining flow
$\Gamma_{\mathrm{UV}} \to \Gamma^*$ integrates out UV
fluctuations while preserving the topological invariant
$J = 1/2$; the residual complexity at the fixed point is the
irreducible cost of maintaining that invariant against vacuum
noise.

%==========================================================================
\section{Majorana Zero Modes and Topological Quantum Computing}
\label{sec:majorana}
%==========================================================================

\subsection{MZMs as Torsion Defects}

In homological codes (Kitaev chain, Toric code), a Majorana Zero Mode
(MZM) acts as a defect at the end of a spin chain. Braiding two MZMs
performs a unitary transformation on the ground state---this ``twist'' is
the topological equivalent of geometric torsion in a 3D spin network.

\subsection{The Proton as a Majorana Bound State}

Viewing the proton's $uud$ structure through a homological lens:
\begin{itemize}
    \item Quarks are modeled as Majorana-like excitations at the nodes
    of the tensor network.
    \item The spin crisis is resolved: spin is stored topologically in
    the strings (gluons) connecting the MZMs.
    \item The Sivers effect becomes the Berry phase acquired by dragging
    a Majorana defect through the network during DIS scattering.
\end{itemize}

\subsection{Wormholes via Majorana Entanglement}

Using the Sachdev--Ye--Kitaev (SYK) model of MZMs in a random network:
\begin{enumerate}
    \item Entangle a pair of MZMs across two separate homological codes
    (representing two ``protons'').
    \item The logical qubit now exists in both locations simultaneously.
    \item The SYK model is holographically dual to a traversable wormhole
    in $\mathrm{AdS}_2$ gravity.
\end{enumerate}

%==========================================================================
\section{Functorial Correspondence: MZMs, Day Convolution, and
Poincar\'{e} Complexes}\label{sec:functor}
%==========================================================================

\subsection{From MZMs to Day Convolution}

A Majorana Zero Mode is an object in a \emph{Modular Tensor Category}
(MTC). We define a functor $\mathcal{F}$ mapping the braiding and fusion
of MZMs into a monoidal category. The \emph{Day convolution}
$\otimes_{\mathrm{Day}}$ acts as the mechanism for merging or
coarse-graining the tensor network---it propagates local spin information
(MZM twists) into global network properties and serves as the engine of
the renormalization group. (The formal definition of Day convolution and
the proof of its uniqueness as the free monoidal completion of the RG
scale poset are given in \cref{sec:day-formal}.)

\subsection{The Poincar\'{e} Complex in Blockchain Cohomology}

Treating spin nodes as a simplicial complex:
\begin{itemize}
    \item \textbf{Blockchain cohomology} treats ``blocks'' (or spin nodes)
    as a simplicial complex. Validating a transaction is equivalent to
    checking that a boundary operator $\partial$ vanishes (the chain is a
    cycle).
    \item A \textbf{Poincar\'{e} complex} satisfies Poincar\'{e} duality:
    the internal spin structure ($k$-cells) perfectly matches the external
    boundary ($(n{-}k)$-cells). It represents a state of
    \emph{topological consensus}.
    \item The functor $\mathcal{F}$ takes the noisy local Majorana braids
    (quarks) and, through the Day convolution, outputs a Poincar\'{e}
    complex---the ``immutable ledger'' of the proton.
\end{itemize}

\subsection{The Wormhole as a Cross-Chain Bridge}

Programming a wormhole is equivalent to constructing a \emph{cross-chain
functor}: the Day convolution ensures that the Poincar\'{e} complex of
Proton~A is entangled (isomorphic) with that of Proton~B. By programming
MZM torsion (the Sivers effect), one writes a ``smart contract'' into the
fabric of spacetime, forcing two distinct regions into a single
homological identity.

%==========================================================================
\section{Boundary Operators and the Sivers Function}\label{sec:boundary}
%==========================================================================

\subsection{The Chain Complex of the Proton}

Define the proton structure as a chain complex:
\begin{equation}\label{eq:chain-complex}
    \cdots
    \xrightarrow{\;\partial_{n+1}\;}
    C_n
    \xrightarrow{\;\partial_n\;}
    C_{n-1}
    \xrightarrow{\;\partial_{n-1}\;}
    \cdots
    \xrightarrow{\;\partial_1\;}
    C_0 \,,
\end{equation}
where $C_0$ (nodes) represents the quarks and $C_1$ (edges) represents the
gluon flux tubes. In a stable proton (a closed chain),
$\partial_n \circ \partial_{n+1} = 0$.

\subsection{The Sivers Function as a Co-boundary Obstruction}

Let $\delta$ be the co-boundary operator (adjoint of $\partial$). The
Sivers function $f_{1T}^\perp$ corresponds to a 1-cocycle in
$H^1$---a characteristic class measuring how the spin bundle is twisted
over the momentum base:
\begin{equation}\label{eq:sivers-cocycle}
    \langle\text{Sivers}\rangle
    \;\propto\;
    \oint_{\partial\Sigma} \omega_{\text{spin}} \,,
\end{equation}
where $\partial\Sigma$ is the boundary of the transverse momentum space.

\subsection{Blockchain Cohomology Interpretation}

\begin{table}[ht]
\centering
\renewcommand{\arraystretch}{1.3}
\begin{tabular}{@{}lll@{}}
\toprule
\textbf{Particle Physics} & \textbf{Categorical Analog} & \textbf{Boundary Math} \\
\midrule
Quark transverse momentum & Local transaction payload & Chain element $c \in C_n$ \\
Proton spin state          & Global ledger consensus   & Cycle $z \in \ker(\partial)$ \\
Sivers function            & Boundary obstruction      & $\delta\omega \neq 0$ \\
OAM                        & Homological loop          & Non-trivial $H_1$ class \\
Gluon field                & Distributed network graph & The simplicial complex \\
\bottomrule
\end{tabular}
\caption{Correspondence between particle physics and cohomological concepts.}
\label{tab:boundary-correspondence}
\end{table}

%==========================================================================
\section{The Poincar\'{e} Protocol and the Spin Crisis as Latency}
\label{sec:latency}
%==========================================================================

The chain complex of \cref{sec:boundary} and its blockchain-cohomology
interpretation lead to a striking re-reading of the proton spin crisis:
the ``missing'' spin is not missing at all; it is unresolved data in the
mempool of the Poincar\'{e} protocol.

\subsection{The Poincar\'{e} Protocol}

\begin{definition}[Poincar\'{e} Protocol]\label{def:poincare-protocol}
The \emph{Poincar\'{e} protocol} is the consensus mechanism by which the
proton's spin network maintains its global topological invariant
$J = 1/2$. It operates on the chain complex
$C_\bullet = (C_n \xrightarrow{\partial_n} C_{n-1} \xrightarrow{} \cdots
\xrightarrow{\partial_1} C_0)$
and enforces two conditions at every energy scale $Q^2$:
\begin{enumerate}
    \item \textbf{Closure.} $\partial \circ \partial = 0$: the boundary
    of a boundary is zero. This guarantees that the total spin ledger is
    self-consistent---no angular momentum is created or destroyed, only
    redistributed among cells.
    \item \textbf{Poincar\'{e} duality.} The $k$-th homology group
    $H_k$ of the complex is isomorphic to the $(n{-}k)$-th cohomology
    group $H^{n-k}$. This enforces a symmetry between the ``interior''
    degrees of freedom (quarks, nodes) and the ``boundary'' degrees of
    freedom (gluons, edges): every unit of spin committed to a node is
    mirrored by a dual unit of spin stored in the surrounding edges.
\end{enumerate}
The protocol is \emph{scale-dependent}: the Day convolution
$\otimes_{\mathrm{Day}}$ acts as the propagation mechanism that resolves
finer structure as the probe energy $Q^2$ increases, analogous to
increasing the block height in a distributed ledger.
\end{definition}

\begin{remark}[Cylindrical consistency as monad associativity]
\label{rem:monad-associativity}
The cylindrical consistency condition
(Eq.~\eqref{eq:cylindrical-consistency}) will be shown in
\cref{sec:syndrome-monad} to be equivalent to the \emph{associativity
law} of the Syndrome Monad:
$\mu \circ \mathsf{T}(\mu) = \mu \circ \mu_{\mathsf{T}}$.
The Poincar\'{e} protocol's requirement that coarse-graining commutes
with embedding is precisely the statement that the RG
decoder $\mu$ is associative---the order in which nested syndrome
layers are collapsed does not matter.
\end{remark}

\subsection{The Wilson Line as a Hash Function}\label{sec:wilson-hash}

The dictionary in \cref{tab:dictionary} identifies the Wilson line as
the ``hash function'' of the Poincar\'{e} protocol. This identification
is not metaphorical; the Wilson line satisfies the defining properties
of a cryptographic hash, with gauge invariance playing the role of
collision resistance.

\subsubsection{Construction}

A quark traversing a path $\gamma$ through the gluon field $A_\mu(x)$
accumulates a net color rotation given by the path-ordered exponential:
\begin{equation}\label{eq:wilson-line}
    W[\gamma]
    = \mathcal{P}\exp\!\left(
        -ig \int_\gamma A_\mu(x)\,dx^\mu
    \right) \;\in\; \mathrm{SU}(3) \,.
\end{equation}
This operator takes the full continuous gauge-field configuration along
$\gamma$---infinitely many degrees of freedom---and compresses it into a
single $\mathrm{SU}(3)$ matrix (8 real parameters).

\subsubsection{Property-by-Property Mapping}

\begin{table}[ht]
\centering
\renewcommand{\arraystretch}{1.3}
\begin{tabular}{@{}p{3.0cm}p{5.0cm}p{5.2cm}@{}}
\toprule
\textbf{Hash Property} &
\textbf{Cryptographic Meaning} &
\textbf{Wilson-Line Realization} \\
\midrule
Deterministic &
Same input $\Rightarrow$ same digest &
Same gauge field $A_\mu$ and path $\gamma$ always yield the same
$\mathrm{SU}(3)$ matrix $W[\gamma]$ \\
Compressive &
Output is much smaller than input &
$\infty$ field degrees of freedom $\to$ 8 real parameters \\
One-way &
Cannot reconstruct input from digest &
Given $W[\gamma]$ alone, the gauge-field configuration along $\gamma$
cannot be recovered; many configurations produce the same $W$
(gauge orbits) \\
Collision resistance &
Hard to find two inputs with the same digest &
Gauge invariance: physically distinct configurations are distinguished
by $W$; gauge-equivalent ones (which are \emph{not} physically
distinct) map to the same orbit \\
Composable &
Hashes can be chained:
$H(H(x)\|y)$ &
Wilson lines compose under path concatenation:
$W[\gamma_1 \circ \gamma_2] = W[\gamma_1]\,W[\gamma_2]$ \\
\bottomrule
\end{tabular}
\caption{Property-by-property correspondence between cryptographic
hash functions and Wilson lines.}
\label{tab:hash-properties}
\end{table}

\subsubsection{What ``Hashing Local Spin Data'' Means Physically}

The phrase ``the Day convolution hashes local spin data into the global
Poincar\'{e} complex'' refers to a three-step physical process:

\begin{enumerate}
    \item \textbf{Local data (the transaction).}
    A quark at node $A$ in the spin network has a definite
    spin-polarization state---say, helicity $+1/2$. In the chain
    complex, this is a chain element $c \in C_0$: a transaction
    waiting to be committed.

    \item \textbf{Hash / propagation (the Wilson line).}
    The gluon flux tube connecting node $A$ to node $B$ defines a
    path $\gamma$. The Wilson line $W[\gamma]$ acts on the quark's
    color-spin state, rotating and mixing its degrees of freedom via
    Eq.~\eqref{eq:wilson-line}. The detailed structure of the gluon
    field along the tube is irrelevant---only the net $\mathrm{SU}(3)$
    rotation (the hash digest) matters. This is gauge invariance in
    action: the ``raw data'' of the gauge field is compressed into a
    single group element.

    \item \textbf{Commitment to the ledger (the global invariant).}
    The quark's spin state \emph{after} rotation by the Wilson line is
    what gets committed to the global Poincar\'{e} complex. The total
    proton spin $J = 1/2$ is the sum over all such committed states
    across all nodes and edges of the network.
\end{enumerate}

\subsubsection{Coarse-Graining as Iterated Hashing}

The RG flow from high $Q^2$ to low $Q^2$ is an \emph{iterated hashing}
process (the formal coarse-graining operations are defined in
\cref{sec:coarse-graining}). At high $Q^2$, the spin network has many nodes (sea quarks)
and edges (gluons), each with their own spin data. As $Q^2$ decreases,
the Day convolution contracts subgraphs into single effective nodes:
\begin{equation}
    \underbrace{
        \bigl\{q_1, q_2, \bar{q}_3, g_4, g_5, \ldots\bigr\}
    }_{\text{high-}Q^2\text{ subgraph}}
    \;\xrightarrow{\;\;W[\gamma_{\text{eff}}]\;\;}\;
    \underbrace{
        q_{\text{valence}}
    }_{\text{low-}Q^2\text{ node}}
    \quad\text{with}\quad
    \Delta q_{\text{valence}}
    = \operatorname{tr}\!\bigl(
        W[\gamma_{\text{eff}}]\;\rho_{\text{subgraph}}
    \bigr) \,.
\end{equation}
A single valence quark at low $Q^2$ is the hash digest of an entire
subgraph at high $Q^2$. The valence quark spin
$\Delta q \approx 0.30$ encodes the net spin of a complex subsystem in
a single number, and one cannot recover the detailed internal structure
from that number alone without probing at higher $Q^2$---exactly the
one-way property of a hash.

This is why the ``spin crisis'' is a latency problem: the valence-quark
spin $\Delta\Sigma$ is a low-resolution hash digest. It faithfully
encodes the \emph{net} spin contribution at that scale, but it has
compressed away the detailed decomposition into $\Delta G$, $L_q$, and
$L_g$. To see those contributions, one must increase $Q^2$ (raise the
block height), forcing the Day convolution to resolve the subgraph and
expose the mempool.

\subsection{Dictionary: Blockchain Cohomology and Particle Physics}

The following table provides a comprehensive dictionary between the
concepts of distributed-ledger consensus and the internal dynamics of
the proton. The left column defines the blockchain concept; the right
column identifies its physical realization.

\begin{longtable}{@{}p{3.8cm}p{3.2cm}p{6.2cm}@{}}
\caption{Complete dictionary between blockchain cohomology and proton
physics.}
\label{tab:dictionary} \\
\toprule
\textbf{Blockchain Concept} &
\textbf{Physics Concept} &
\textbf{Formal Identification} \\
\midrule
\endfirsthead
\toprule
\textbf{Blockchain Concept} &
\textbf{Physics Concept} &
\textbf{Formal Identification} \\
\midrule
\endhead
\midrule
\multicolumn{3}{r}{\itshape Continued on next page} \\
\bottomrule
\endfoot
\bottomrule
\endlastfoot

% --- Structural ---
Distributed ledger &
Proton spin network &
The simplicial complex $C_\bullet$ whose cells are quarks (nodes)
and gluon flux tubes (edges) \\

Confirmed block &
Valence quark state &
A $0$-cycle $z \in \ker(\partial_1) \subset C_0$: angular momentum
that has been locally committed to a node \\

Mempool (unconfirmed transactions) &
Sea quarks and gluons &
Syndrome data: non-trivial chains in $C_1, C_2, \ldots$ that carry
angular momentum in the edges and higher cells \\

Block height &
Probe resolution $Q^2$ &
The energy scale at which the Day convolution has propagated; higher
$Q^2$ resolves finer cells in the complex \\

Global ledger state &
Total proton spin $J = 1/2$ &
The homology class $[z] \in H_0(C_\bullet)$: the logical qubit,
always exactly $1/2$ regardless of block height \\

% --- Consensus ---
Consensus protocol &
Poincar\'{e} protocol &
The closure condition $\partial \circ \partial = 0$ plus Poincar\'{e}
duality (Def.~\ref{def:poincare-protocol}) \\

Mining / validation &
Day convolution $\otimes_{\mathrm{Day}}$ &
The functorial operation that hashes local spin data (MZM braids)
into the global Poincar\'{e} complex \\

Hash function &
Wilson line &
The path-ordered exponential of the gauge field; maps a quark path
(transaction route) to a color/spin state (hash digest) \\

Proof of work &
Termination proof &
\texttt{WellFounded.fromMeasure}: a decreasing measure guaranteeing
that the consensus computation halts \\

% --- Dynamics ---
Transaction &
Quark momentum transfer &
A chain element $c \in C_n$ representing a parton carrying transverse
momentum $\vb{k}_\perp$ \\

Transaction fee (gas) &
Torsional phase $\phi$ &
The Sivers phase: the ``cost'' of routing a transaction through the
network; determines the direction of the asymmetry \\

Fork &
Torsion / Sivers asymmetry &
A non-trivial element in the torsion subgroup of $H^n(X; \mathbb{Z})$;
the network branches when spin is polarized \\

Fork resolution &
Syzygy validation &
The requirement that any fork be resolved into a valid Poincar\'{e}
complex within $t_{\text{Planck}}$ cycles \\

Double-spend attack &
Spin violation $J \neq 1/2$ &
Topologically forbidden: the Poincar\'{e} protocol guarantees
$[z] = 1/2$ at all scales \\

% --- Latency ---
Latency &
RG flow / TMD evolution &
The ``time'' (in resolution-scale) required for the Day convolution to
propagate spin data from the mempool into the confirmed blocks \\

Mempool backlog &
``Missing'' spin ($70\%$) &
Angular momentum stored in $C_1$ (gluon spin $\Delta G$) and in
non-trivial $H_1$ classes (orbital angular momentum $L_q + L_g$) \\

Confirmation depth &
Number of RG steps &
How many coarse-graining iterations (Pachner moves) have been applied;
deeper confirmation = lower $Q^2$ = fewer resolved degrees of freedom \\

% --- Error correction ---
Syndrome &
Sea quark/gluon excitation &
A stabilizer eigenvalue $-1$; a quasi-particle generated when the
probe energy exceeds the code's gap \\

Error correction &
Confinement &
The decoder contracts the syndrome cloud back into a color-singlet
(the logical qubit $J = 1/2$) \\

Logical qubit &
Baryon number / total spin &
The topological invariant protected by the homological code; immune to
local perturbations (syndrome noise) \\

Code distance $d$ &
Confinement scale $\Lambda_{\text{QCD}}$ &
The minimum number of syndrome errors required to corrupt the logical
state; $d \sim 1/\Lambda_{\text{QCD}}$ \\

% --- Cross-chain ---
Cross-chain bridge &
Wormhole (ER bridge) &
Entangled Poincar\'{e} complexes: the Day convolution connects two
protons into a shared syndrome space \\

Smart contract &
Topological transfer protocol &
The \texttt{WormholeBridgeApp} of \cref{sec:smart-contract}: a
self-executing set of constraints governing syndrome teleportation \\
\end{longtable}

\subsection{The Spin Crisis as a Latency Problem}

With the dictionary in hand, the proton spin crisis admits a precise
re-statement.

\subsubsection{The Standard Framing}

In QCD, the spin sum rule reads
\begin{equation}\label{eq:spin-sum-redux}
    J = \frac{1}{2}
      = \underbrace{\tfrac{1}{2}\Delta\Sigma}_{\text{quark spin}}
      + \underbrace{\Delta G}_{\text{gluon spin}}
      + \underbrace{L_q + L_g}_{\text{orbital angular momentum}} \,.
\end{equation}
A DIS experiment at resolution $Q^2$ measures $\Delta\Sigma$ and finds
it accounts for only $\sim\!30\%$ of $1/2$. The ``crisis'' assumes that
all the spin \emph{should} reside at the quarks and asks: where did the
rest go?

\subsubsection{The Blockchain Cohomology Reframing}

In the Poincar\'{e} protocol, the total spin $J = 1/2$ is the
\emph{logical qubit}---the global homology class $[z] \in H_0$.
It is \emph{always} exactly $1/2$, just as the confirmed state of a
well-formed blockchain is always self-consistent. There is no missing
spin, and there never was.

The apparent deficit arises from \emph{what we measure and when}:

\begin{description}[leftmargin=0pt, style=nextline]
    \item[Valence quarks are confirmed blocks.]
    They are the coarse-grained nodes of the spin network---the data
    that has been hashed into the global consensus at the scale $Q^2$
    of the probe. They sit in $C_0$ and carry $\sim\!30\%$ of the spin.
    This is not a failure; it is the amount of angular momentum that has
    been \emph{locally committed} to the node level of the chain
    complex.

    \item[Sea quarks and gluons are the mempool.]
    They are the syndromes---the edges ($C_1$) and higher cells of
    the complex. They carry real angular momentum ($\Delta G$ and
    $L_q + L_g$), but this information is distributed across the
    \emph{connections} of the network, not localized at the nodes.
    These are valid transactions that exist in the network but have not
    yet been mined into a block at the resolution being queried.

    \item[The Day convolution is the propagation protocol.]
    It is the mechanism by which local spin information (a braid twist
    at one MZM node) gets hashed into the global Poincar\'{e} complex.
    This propagation takes ``time''---not clock time, but
    \emph{resolution time}. As $Q^2$ increases, the Day convolution
    resolves finer structure, and more of the mempool becomes visible.

    \item[The ``latency'' is the RG flow.]
    The renormalization group flow from high $Q^2$ to low $Q^2$ is the
    coarse-graining (syndrome decoding) that contracts the full tensor
    network into the effective valence picture. Information is not
    physically lost but is \emph{not yet resolved} at the scale of the
    probe. The spin ``deficit'' is precisely the information that has
    been coarse-grained into the edges and is waiting to be resolved at
    a finer scale.
\end{description}

The crisis, then, is a \emph{measurement artifact}: we queried the
ledger at a coarse block height and were surprised that not all
transactions had been confirmed. The total is always $1/2$. The question
was never ``where did the spin go?'' but ``at what resolution does each
contribution become visible?''

\subsubsection{The DGLAP Running as Mempool Dynamics}

This interpretation also explains a specific experimental fact: under
DGLAP evolution, the quark spin fraction $\Delta\Sigma$ \emph{decreases}
logarithmically as $Q^2$ increases, while the gluon contribution
$\Delta G$ grows. In the blockchain picture, this is exactly what one
expects. As resolution increases, transactions migrate from the
``confirmed node'' category into the newly resolved ``edge/syndrome''
category, because higher resolution reveals that what looked like a
simple node is actually a subgraph with its own internal structure.

Formally, let $N_{\text{confirmed}}(Q^2)$ denote the number of
confirmed blocks (valence degrees of freedom) and
$N_{\text{mempool}}(Q^2)$ the mempool size (sea quarks + gluons) at
scale $Q^2$. The DGLAP splitting functions $P_{qq}$, $P_{qg}$,
$P_{gq}$, $P_{gg}$ are the \emph{transaction-routing rules} of the
Poincar\'{e} protocol: they govern how angular momentum is transferred
between nodes and edges as the Day convolution resolves one scale step.
The Altarelli--Parisi equation
\begin{equation}\label{eq:dglap-latency}
    \frac{\partial \Delta q(x, Q^2)}{\partial \ln Q^2}
    = \frac{\alpha_s}{2\pi}
      \int_x^1 \frac{dy}{y}\,
      \left[
          \Delta P_{qq}\!\left(\frac{x}{y}\right) \Delta q(y, Q^2)
        + \Delta P_{qg}\!\left(\frac{x}{y}\right) \Delta g(y, Q^2)
      \right]
\end{equation}
describes how the confirmed-block content $\Delta q$ evolves as the
block height $\ln Q^2$ increases: quarks split into quark--gluon pairs
($P_{qg}$), moving angular momentum from nodes to edges, while gluons
split into quark--antiquark pairs ($P_{gq}$), promoting mempool data
into newly created nodes.

\subsubsection{Mempool Anatomy: Three Compartments of Hidden Spin}

The mempool of the Poincar\'{e} protocol is not a structureless pool.
It decomposes into three geometrically distinct compartments, each
carrying a different species of angular momentum. This decomposition
arises naturally from the chain complex $C_2 \xrightarrow{\partial_2}
C_1 \xrightarrow{\partial_1} C_0$ that defines the proton's
Poincar\'{e} complex (see \cref{sec:latency}).

\paragraph{Compartment 1: Gluon spin $\Delta G$ in edges ($C_1$).}
Gluon spin resides in the \emph{edges} of the chain complex---the
connections between nodes. In the gauge theory, the gluon field
$A_\mu^a$ is itself a connection on a principal $\mathrm{SU}(3)$
bundle; its spin polarization $\Delta G$ is the angular momentum stored
in these connections rather than at the nodes. The blockchain analog is
\emph{value in transit}: funds that have left one account but have not
yet been credited to another.  They are real (conserved), but invisible
to any single-node query. Formally,
\begin{equation}\label{eq:gluon-spin-edge}
    \Delta G(Q^2)
    = \sum_{e \in C_1(Q^2)} \sigma(e)\,,
\end{equation}
where $\sigma(e)$ is the helicity weight carried by edge $e$ at
resolution $Q^2$.

\paragraph{Compartment 2: Quark OAM $L_q$ in 1-cycles ($H_1$).}
Quark orbital angular momentum lives not in individual edges but in
\emph{non-trivial closed loops}---elements of the first homology group
$H_1 = \ker\partial_1 / \operatorname{im}\partial_2$. A quark orbiting
the proton's center traces a closed path in the chain complex; its
winding number around non-contractible cycles contributes to $L_q$.
The blockchain analog is a \emph{circular transaction chain}: a cycle
$A \to B \to C \to A$ whose net winding number is non-zero. Such
cycles are closed ($\partial_1 = 0$) but not exact (not the boundary of
a 2-cell), so they represent genuine topological structure:
\begin{equation}\label{eq:quark-oam-h1}
    L_q(Q^2)
    = \sum_{[\gamma] \in H_1(Q^2)}
      \operatorname{wind}(\gamma) \cdot |\gamma|\,,
\end{equation}
where $\operatorname{wind}(\gamma)$ is the winding number and
$|\gamma|$ the cycle's angular-momentum magnitude. Ji's sum rule
\cite{Ji1997} provides experimental access via generalized parton
distributions (GPDs):
$J_q = \tfrac{1}{2}\bigl[A_{q}(0) + B_{q}(0)\bigr]$,
and $L_q = J_q - \tfrac{1}{2}\Delta\Sigma_q$.

\paragraph{Compartment 3: Gluon OAM $L_g$ in 2-cycles ($H_2$).}
Gluon orbital angular momentum occupies the deepest topological
stratum: the \emph{second homology group}
$H_2 = \ker\partial_2 / \operatorname{im}\partial_3$. These are field
configurations that wind around the proton as closed surfaces---membrane
modes in the Poincar\'{e} complex. They are the most ``latent'' of all
spin contributions, requiring the highest resolution to detect. The
blockchain analog is a \emph{self-referential smart-contract loop}: a
protocol-level construct that generates circular flow without any
individual transaction being circular. Formally,
\begin{equation}\label{eq:gluon-oam-h2}
    L_g(Q^2)
    = \sum_{[\Sigma] \in H_2(Q^2)}
      \operatorname{wind}(\Sigma) \cdot |\Sigma|\,,
\end{equation}
where the sum runs over non-trivial 2-cycles $\Sigma$.

\paragraph{Resolution hierarchy.}
The three compartments become experimentally accessible at different
$Q^2$ scales, exactly as a blockchain explorer reveals deeper structure
as it indexes more of the network:

\begin{table}[h]
\centering
\caption{Resolution hierarchy of mempool compartments.}
\label{tab:resolution-hierarchy}
\begin{tabular}{@{}llll@{}}
\toprule
\textbf{Scale $Q^2$} & \textbf{Visible compartment}
  & \textbf{Experiment} & \textbf{Blockchain analog} \\
\midrule
$\sim 1\;\text{GeV}^2$
  & Valence quarks ($C_0$)
  & DIS (EMC, HERMES)
  & Confirmed blocks \\
$\sim 10\;\text{GeV}^2$
  & $+\;\Delta G$ in edges ($C_1$)
  & Polarized $pp$ (RHIC/STAR)
  & $+\;$In-transit value \\
$\sim 10\text{--}100\;\text{GeV}^2$
  & $+\;L_q$ in 1-cycles ($H_1$)
  & DVCS / GPDs (JLab, EIC)
  & $+\;$Circular chains \\
$\gtrsim 100\;\text{GeV}^2$
  & $+\;L_g$ in 2-cycles ($H_2$)
  & Diffractive dijets (EIC)
  & $+\;$Protocol loops \\
\bottomrule
\end{tabular}
\end{table}

\noindent
At each step, the Day convolution resolves one more topological
stratum of the mempool, moving angular momentum from ``unconfirmed''
to ``visible.'' The spin sum rule $\frac{1}{2} = \frac{1}{2}
\Delta\Sigma + \Delta G + L_q + L_g$ holds at every scale, but
the partition among terms shifts as the indexing depth increases.
This is why the EMC experiment at $Q^2 \sim 10\;\text{GeV}^2$ saw
only $\sim\!30\%$ of the spin in quark helicities: the 1-cycle and
2-cycle compartments ($L_q$ and $L_g$) were still deep in the
mempool, awaiting confirmation by higher-resolution probes.

\subsubsection{The Asymptotic Limit}

At $Q^2 \to \infty$ (infinite resolution), the Day convolution has
fully propagated, the entire mempool has been confirmed, and the
complete spin budget is visible:
\begin{equation}
    \lim_{Q^2 \to \infty}
    \left[
        \tfrac{1}{2}\Delta\Sigma(Q^2)
        + \Delta G(Q^2)
        + L_q(Q^2) + L_g(Q^2)
    \right]
    = \frac{1}{2} \,.
\end{equation}
This is the blockchain reaching \emph{finality}: every transaction has
been confirmed, every fork has been resolved, and the ledger agrees
with the logical qubit. At any finite $Q^2$, the sum still equals
$1/2$ (the sum rule is exact), but the \emph{partition} among the
four terms depends on how much of the network has been resolved.
The ``crisis'' is the observation that, at the $Q^2$ values accessible
to the EMC experiment ($\sim\!10\;\text{GeV}^2$), most of the angular
momentum was still in the mempool.

%==========================================================================
\section{The Syndrome Monad}\label{sec:syndrome-monad}
%==========================================================================

The preceding sections have implicitly used three categorical
structures: (i)~syndrome generation as a computational side effect,
(ii)~Day convolution as the mechanism for propagating information across
scales, and (iii)~the MWPM decoder as a coarse-graining operation.
We now unify all three into a single algebraic object---a
\emph{monad}~\cite{Moggi1991,Riehl2016}---which provides the
mathematical backbone for every construction that follows.

\subsection{Definition of the Syndrome Monad}

\begin{definition}[The category $\mathcal{C}$]
\label{def:syndrome-category}
Let $\mathcal{C}$ be the category whose objects are graded chain
complexes $(C_\bullet, \partial)$ over a compact oriented 3-manifold
$\Sigma$ and whose morphisms are chain maps (degree-preserving linear
maps commuting with $\partial$).
\end{definition}

\begin{definition}[Syndrome Monad]
\label{def:syndrome-monad}
The \emph{Syndrome Monad} is the triple
$(\mathsf{T}, \eta, \mu)$ defined as follows.

\paragraph{Endofunctor $\mathsf{T} : \mathcal{C} \to \mathcal{C}$.}
On objects:
\begin{equation}\label{eq:syndrome-functor}
    \mathsf{T}(C_\bullet)_k
    \;=\;
    C_k \;\oplus\; \mathrm{Synd}_k(C_\bullet) \,,
\end{equation}
where $\mathrm{Synd}_k(C_\bullet)$ is the syndrome space at chain
degree $k$---the space of stabilizer violations of the boundary
operator $\partial$ at rank $k$. Physically:
\begin{itemize}
    \item $k = 0$: $\mathrm{Synd}_0$ = sea quark excitations (primal /
    $Z$-type syndromes; cf.\ \cref{sec:syndromes}).
    \item $k = 1$: $\mathrm{Synd}_1$ = gluon excitations (dual /
    $X$-type syndromes).
    \item $k = 2$: $\mathrm{Synd}_2$ = membrane / vacuum fluctuations.
\end{itemize}
On morphisms: a chain map $f : C_\bullet \to D_\bullet$ lifts to
$\mathsf{T}(f) = f \oplus f^{\mathrm{synd}}$, where
$f^{\mathrm{synd}}$ is the induced map on syndrome spaces. This is
well-defined and covariant because syndromes are defined by $\partial$,
which chain maps preserve.

\paragraph{Unit $\eta : \mathrm{Id} \Rightarrow \mathsf{T}$.}
The canonical inclusion into the first summand:
\begin{equation}\label{eq:monad-unit}
    \eta_{C_\bullet} : C_k \;\hookrightarrow\;
    C_k \oplus \mathrm{Synd}_k(C_\bullet) \,,
    \qquad
    c \;\mapsto\; (c, 0) \,.
\end{equation}
Physical meaning: \emph{state initialization}---the bare proton with
zero syndromes.

\paragraph{Multiplication $\mu : \mathsf{T}^2 \Rightarrow \mathsf{T}$.}
Collapses nested syndrome layers (``syndromes of syndromes'') via the
torsional MWPM decoder (\cref{sec:decoder}):
\begin{equation}\label{eq:monad-mult}
    \mu_{C_\bullet}
    : \mathsf{T}\bigl(\mathsf{T}(C_\bullet)\bigr)_k
    \;=\;
    (C_k \oplus \mathrm{Synd}_k)
    \oplus
    \mathrm{Synd}_k(C_\bullet \oplus \mathrm{Synd}_\bullet)
    \;\longrightarrow\;
    C_k \oplus \mathrm{Synd}_k \,.
\end{equation}
Physical meaning: \emph{the RG step}---coarse-graining nested quantum
corrections into a single effective correction.
\end{definition}

\subsection{Monad Laws}

\begin{proposition}[Monad laws for $\mathsf{T}$]
\label{prop:monad-laws}
The triple $(\mathsf{T}, \eta, \mu)$ satisfies the three monad laws:
\begin{enumerate}
    \item \textbf{Left unit.}
    $\mu \circ \mathsf{T}(\eta) = \mathrm{id}$:
    decoding a freshly initialized syndrome layer (one that carries
    zero syndromes) returns the original state.

    \item \textbf{Right unit.}
    $\mu \circ \eta_{\mathsf{T}} = \mathrm{id}$:
    embedding a syndromic state into $\mathsf{T}^2$ as a ``clean
    outer layer'' and then decoding recovers the original.

    \item \textbf{Associativity.}
    $\mu \circ \mathsf{T}(\mu)
    = \mu \circ \mu_{\mathsf{T}}$:
    RG flow is \emph{path-independent}. This is precisely the
    cylindrical consistency condition
    (\cref{def:cylindrical-consistency},
    Eq.~\eqref{eq:cylindrical-consistency}).
\end{enumerate}
The laws are summarized by the commutative diagrams:
\[
\begin{tikzcd}
    \mathsf{T}^3 \ar[r, "\mathsf{T}(\mu)"] \ar[d, "\mu_{\mathsf{T}}"']
    & \mathsf{T}^2 \ar[d, "\mu"] \\
    \mathsf{T}^2 \ar[r, "\mu"']
    & \mathsf{T}
\end{tikzcd}
\qquad
\begin{tikzcd}
    \mathsf{T} \ar[r, "\mathsf{T}(\eta)"] \ar[dr, "\mathrm{id}"']
    & \mathsf{T}^2 \ar[d, "\mu"]
    & \mathsf{T} \ar[l, "\eta_{\mathsf{T}}"'] \ar[dl, "\mathrm{id}"] \\
    & \mathsf{T}
\end{tikzcd}
\]
\end{proposition}

\subsection{The Kleisli Category = Physical State Transitions}

\begin{proposition}[Kleisli category of the Syndrome Monad]
\label{prop:kleisli}
The Kleisli category $\mathrm{Kl}(\mathsf{T})$ has:
\begin{itemize}
    \item \textbf{Objects}: chain complexes $C_\bullet$
    (= proton states at a given resolution).
    \item \textbf{Morphisms} $A \to B$: maps
    $f : A \to \mathsf{T}(B)$, i.e.\ state transitions that
    \emph{may generate syndromes}.
    \item \textbf{Identity}: $\eta_A : A \to \mathsf{T}(A)$
    (the ``do nothing'' transition).
    \item \textbf{Composition}: for $f : A \to \mathsf{T}(B)$ and
    $g : B \to \mathsf{T}(C)$,
    \begin{equation}\label{eq:kleisli-comp}
        (g \mathbin{>\!\!=\!\!>} f)(a)
        \;=\;
        \mu\bigl(\mathsf{T}(g)(f(a))\bigr) \,.
    \end{equation}
\end{itemize}
\end{proposition}

The physical interpretation is immediate:
\begin{itemize}
    \item A DIS collision is a Kleisli morphism: it takes a proton
    state to a final state \emph{plus} syndrome excitations.
    \item Kleisli composition $(g \mathbin{>\!\!=\!\!>} f)$ is the
    \texttt{>\!>\!>} operator in the Babel DSL
    (\cref{sec:smart-contract}).
    \item In the language of Moggi~\cite{Moggi1991}: syndrome
    generation is the \emph{computational side effect}; pure
    computations (via $\eta$) are topological operations that preserve
    the code state exactly.
\end{itemize}

\subsection{Rank-Graded Decomposition}\label{sec:rank-graded}

The Syndrome Monad decomposes by chain complex rank:
\begin{equation}\label{eq:rank-decomp}
    \mathsf{T}
    \;=\;
    \mathsf{T}_0 \times \mathsf{T}_1 \times \cdots \times \mathsf{T}_n \,,
\end{equation}
where $\mathsf{T}_k$ acts on the degree-$k$ component. The boundary
operator $\partial_k : C_k \to C_{k-1}$ induces a natural
transformation
$\partial_k^* : \mathsf{T}_k \Rightarrow \mathsf{T}_{k-1}$.

\begin{definition}[Promotion functor]
\label{def:promotion}
The \emph{promotion functor}
$\Phi_k : \mathrm{Kl}(\mathsf{T}_k) \to \mathrm{Kl}(\mathsf{T}_{k+1})$
maps Kleisli morphisms across adjacent ranks. The collection
$\{\Phi_k\}_{k \geq 0}$ assembles into a \emph{chain functor}:
\begin{equation}\label{eq:chain-functor}
    \Phi_{k-1} \circ \partial_k^*
    \;=\;
    \partial_k^* \circ \Phi_k \,.
\end{equation}
\end{definition}

\begin{proposition}[DGLAP as natural transformation]
\label{prop:dglap}
The DGLAP splitting functions $P_{qq}$, $P_{qg}$, $P_{gq}$, $P_{gg}$
are natural transformations between rank-graded monad components:
\begin{equation}\label{eq:dglap-nat}
    P_{ab} : \mathsf{T}_a \Rightarrow \mathsf{T}_b \,,
    \qquad a, b \in \{0, 1\} \,.
\end{equation}
\end{proposition}

\begin{proposition}[Promotion preserves Day convolution]
\label{prop:promotion-day}
The promotion functor is monoidal:
\begin{equation}\label{eq:promotion-monoidal}
    \Phi_k(f \otimes_{\mathrm{Day}} g)
    \;=\;
    \Phi_k(f) \otimes_{\mathrm{Day}} \Phi_k(g) \,.
\end{equation}
\end{proposition}

\subsection{Hierarchical Configuration Spaces}
\label{sec:config-spaces}

The rank-graded structure of $\mathsf{T}$ induces a hierarchy of
configuration spaces built by iterated tensor products.

\begin{definition}[Inductive configuration spaces]
\label{def:config-spaces}
The configuration space at chain degree $k$ is defined inductively:
\begin{align}
    \mathrm{Conf}(C_0)
    &= V_{1/2}^{\mathrm{SU}(2)} \otimes V_3^{\mathrm{SU}(3)}
    \qquad\text{(quark: spin-$1/2$, color triplet)} \,,
    \label{eq:conf-base} \\
    \mathrm{Conf}(C_k)
    &= \mathrm{Conf}(C_{k-1})^{\otimes n_k}
       \big/ \mathrm{Inv}_k
    \qquad\text{(tensor product modulo gauge invariance)} \,,
    \label{eq:conf-inductive}
\end{align}
where $n_k$ is the number of $(k{-}1)$-cells incident on each
$k$-cell and $\mathrm{Inv}_k$ denotes the gauge-invariant subspace.
\end{definition}

This definition is already implicit in the colored spin-network Hilbert
space (Eq.~\eqref{eq:colored-hilbert}): the intertwiner
$\mathcal{I}_v$ at each node is exactly $\mathrm{Inv}$ of the tensor
product of incident edge representations.

For the proton specifically:
\begin{align}
    \mathrm{Conf}(C_0)
    &= V_{1/2} \otimes V_3
    &&\text{(quark)} \,,
    \label{eq:conf-proton-0} \\
    \mathrm{Conf}(C_1)
    &= \mathrm{Conf}(C_0)^{\otimes 2} / \mathrm{Inv}_1
    &&\text{(gluon flux tube: two quarks, projected)} \,,
    \label{eq:conf-proton-1} \\
    \mathrm{Conf}(C_2)
    &= \mathrm{Conf}(C_1)^{\otimes n_2} / \mathrm{Inv}_2
    &&\text{(plaquettes built from edges)} \,.
    \label{eq:conf-proton-2}
\end{align}

\begin{proposition}[Complexity recurrence]
\label{prop:complexity-recurrence}
The computational complexity at rank $k$ satisfies:
\begin{equation}\label{eq:complexity-recurrence}
    C_k = n_k \cdot C_{k-1} - \ln|\mathrm{Inv}_k| \,.
\end{equation}
The gauge invariance projection ($-\ln|\mathrm{Inv}_k|$) represents
the computational savings from symmetry. The exponential growth
($n_k \cdot C_{k-1}$) explains why higher-rank contributions
($L_g$ in $H_2$) are ``most latent.''
\end{proposition}

\begin{proposition}[Extension into $\mathrm{AdS}_5$]
\label{prop:ads-extension}
The chain complex $C_2 \to C_1 \to C_0$ of the proton lives in
$3{+}1$D. In $\mathrm{AdS}_5$, the holographic variable $\zeta$
provides a fifth dimension, and the chain complex extends:
\begin{equation}\label{eq:ads-chain}
    C_3^{\mathrm{AdS}}
    \;\xrightarrow{\;\partial_3\;}
    C_2
    \;\xrightarrow{\;\partial_2\;}
    C_1
    \;\xrightarrow{\;\partial_1\;}
    C_0 \,,
\end{equation}
where $C_3$ encodes the bulk volume cells along the AdS radial
direction. The configuration space at rank~3 is:
\begin{equation}\label{eq:conf-ads}
    \mathrm{Conf}(C_3)
    = \mathrm{Conf}(C_2)^{\otimes n_3} / \mathrm{Inv}_3 \,.
\end{equation}
This gives the tensor-product hierarchy a direct geometric
interpretation: moving deeper into the AdS bulk (larger $\zeta$)
corresponds to ascending the chain complex ranks and tensoring
configuration spaces.
\end{proposition}

\subsection{Day Convolution as State Description}
\label{sec:day-formal}

The paper has used Day convolution informally throughout
(\cref{sec:functor,sec:latency}). We now supply the formal definition
and prove that it is the \emph{unique} monoidal structure compatible
with RG flow.

\begin{definition}[Day convolution]
\label{def:day-convolution}
Let $(\mathcal{P}, \otimes, I)$ be the poset of RG scales $Q^2$,
ordered by flow (higher $Q^2$ is finer resolution). For presheaves
$F, G \in [\mathcal{P}, \mathrm{Vect}]$, the \emph{Day
convolution}~\cite{Day1970} is:
\begin{equation}\label{eq:day-convolution}
    (F \otimes_{\mathrm{Day}} G)(p)
    \;=\;
    \int^{a, b \in \mathcal{P}}
    \mathcal{P}(a \otimes b,\, p)
    \;\otimes\;
    F(a) \otimes G(b) \,,
\end{equation}
where $\int^{a,b}$ denotes the coend---the categorified ``sum over
intermediate scales.''
\end{definition}

For the proton: let $F$ be the quark spin presheaf and $G$ the gluon
spin presheaf. Then:
\begin{equation}\label{eq:day-proton}
    (F \otimes_{\mathrm{Day}} G)(Q^2)
    \;=\;
    \sum_{Q_1^2 \cdot Q_2^2 \leq Q^2}
    F(Q_1^2) \otimes G(Q_2^2) \,.
\end{equation}

\begin{theorem}[Fourier--Day correspondence]
\label{thm:fourier-day}
The Day convolution on $[\mathcal{P}, \mathrm{Vect}]$ is the
categorical analog of the classical Fourier transform. The
correspondence has four legs:
\begin{enumerate}
    \item \textbf{Multiplication $\leftrightarrow$ Convolution.}
    The monoidal product in the scale category $\mathcal{P}$ becomes
    Day convolution in $[\mathcal{P}, \mathrm{Vect}]$.

    \item \textbf{Parseval $\leftrightarrow$ Spin sum rule.}
    The total spin $J = 1/2$ is preserved regardless of the
    resolution basis:
    $\tfrac{1}{2}\Delta\Sigma + \Delta G + L_q + L_g = 1/2$ at every
    $Q^2$.

    \item \textbf{Uncertainty $\leftrightarrow$ Spin crisis.}
    The quark spin uncertainty $\Delta(\Delta\Sigma)$ and gluon spin
    uncertainty $\Delta(\Delta G)$ satisfy a trade-off: measuring
    quark spin precisely leaves gluon spin unresolved.

    \item \textbf{Inverse transform $\leftrightarrow$ Syndrome
    decoding.}
    The monad multiplication $\mu$ recovers the low-resolution state
    from the high-resolution syndrome decomposition.
\end{enumerate}
\end{theorem}

\begin{definition}[Proton state as Day convolution]
\label{def:proton-day}
The proton state at scale $Q^2$ is the Day convolution of rank-graded
presheaves:
\begin{equation}\label{eq:proton-day}
    \Psi_{\mathrm{proton}}(Q^2)
    \;=\;
    \bigl(\eta \otimes_{\mathrm{Day}}
    F_0 \otimes_{\mathrm{Day}}
    F_1 \otimes_{\mathrm{Day}}
    F_2\bigr)(Q^2) \,,
\end{equation}
where $F_k$ is the presheaf of rank-$k$ syndrome data and $\eta$ is
the monad unit (the bare, zero-syndrome initial state). This replaces
the informal statement ``spacetime is a Day convolution of homological
codes'' with a precise mathematical object.
\end{definition}

\begin{theorem}[Uniqueness of Day convolution]
\label{thm:day-unique}
The Day convolution $\otimes_{\mathrm{Day}}$ is the free monoidal
completion of $\mathcal{P}$~\cite{Day1970}. It is therefore the
\emph{unique} monoidal structure on $[\mathcal{P}, \mathrm{Vect}]$
compatible with the RG flow ordering on $\mathcal{P}$. The proton
state description~\eqref{eq:proton-day} is not a modeling choice but
a forced consequence of the categorical structure.
\end{theorem}

\begin{proposition}[$\mathsf{T}$ is a monoidal monad]
\label{prop:monoidal-monad}
The Syndrome Monad is compatible with Day convolution:
\begin{equation}\label{eq:monoidal-monad}
    \mathsf{T}(F \otimes_{\mathrm{Day}} G)
    \;\cong\;
    \mathsf{T}(F) \otimes_{\mathrm{Day}} \mathsf{T}(G) \,.
\end{equation}
Physically: generating syndromes for a composite system is equivalent
to generating syndromes for each component and then convolving.
\end{proposition}

%==========================================================================
\section{Quantum Circuit for Torsional Homology}\label{sec:circuit}
%==========================================================================

\subsection{Architecture}

The circuit is designed for a topological quantum computer (braided MZM
array or high-fidelity surface-code processor). The proton's tensor
network is encoded in a surface code, and ``Sivers-style torsion'' is
injected via a boundary operator linking spin-polarization logic to
transverse spatial displacement.

\subsection{Pseudocode}

\begin{lstlisting}[style=python, caption={Quantum circuit for torsional homology.}]
# HYPER-PARAMETERS
LATTICE_SIZE = (d, d)      # Distance d of the homological code
SIVERS_PHASE = phi         # Torsion angle from Day Convolution
TRANSVERSE_SHIFT = delta_k

# 1. INITIALIZATION: CREATE THE POINCARE COMPLEX
circuit = QuantumCircuit(Lattice)
circuit.initialize_surface_code(LATTICE_SIZE)

# 2. DEFECT CREATION: INJECT MAJORANA ZERO MODES (QUARKS)
MZM_u1 = circuit.create_defect(position=(x1, y1), type="Primal")
MZM_u2 = circuit.create_defect(position=(x2, y2), type="Primal")
MZM_d  = circuit.create_defect(position=(x3, y3), type="Dual")

# 3. TORSION INJECTION: THE SIVERS BOUNDARY OPERATOR
def apply_sivers_torsion(circuit, spin_state, phase):
    """
    Co-boundary Operator: Spin -> Geometric Twist.
    Equivalent to the 'Sign Change' in the Day Convolution.
    """
    if spin_state == "TRANSVERSE_UP":
        circuit.braid(MZM_u1, MZM_d, angle=phase)
    elif spin_state == "TRANSVERSE_DOWN":
        circuit.braid(MZM_u1, MZM_d, angle=-phase)

# 4. TRANSVERSE MOMENTUM SHIFT (DIS "KICK")
circuit.translate_defect(MZM_u1, vector=(TRANSVERSE_SHIFT, 0))

# 5. MEASUREMENT: EXTRACT THE COHOMOLOGY CLASS
syndrome_data = circuit.measure_all_stabilizers()

# 6. POINCARE DUALITY CHECK
if check_consensus(syndrome_data):
    return "Stable Sivers-Active Proton State"
else:
    return "Wormhole Instability / Logical Error"
\end{lstlisting}

\subsection{Mapping to the Mathematics}

\begin{itemize}
    \item \texttt{initialize\_surface\_code}: sets up the Poincar\'{e}
    complex; qubits are $k$-cells, stabilizers are boundary operators
    $\partial$.

    \item \texttt{apply\_sivers\_torsion}: the Day convolution in action;
    applies the spin functor as a topological phase (torsion) across the
    network edges.

    \item The braiding angle $\phi$ is calculated via the AdS/QCD
    light-front holography equations of \cref{sec:lfh}, matching the
    anomalous magnetic moment.

    \item \texttt{translate\_defect}: mimics the Sivers effect where the
    proton spin results in a preferential sideways displacement during
    collision.
\end{itemize}

%==========================================================================
\section{Syndrome Data as Sea Quarks and Gluons}\label{sec:syndromes}
%==========================================================================

\subsection{From Errors to Physics}

In an ideal code, all boundary operators return zero syndromes:
$\partial\ket{\Psi} = 0$. This represents the ``bare'' proton at very low
energy. As probe energy $Q^2$ increases, quasi-particle excitations
``break'' the consensus, generating non-trivial syndromes.

\begin{table}[ht]
\centering
\renewcommand{\arraystretch}{1.3}
\begin{tabular}{@{}lll@{}}
\toprule
\textbf{Syndrome Type} & \textbf{Topological Error} & \textbf{QCD Analog} \\
\midrule
Primal ($Z$-type) & Chain termination & Sea quark/antiquark pair \\
Dual ($X$-type) & Flux-loop excitation & Gluon excitation \\
String operators & Path between syndromes & Gluon flux tube \\
\bottomrule
\end{tabular}
\caption{Syndrome--parton correspondence.}
\label{tab:syndromes}
\end{table}

The spin crisis is resolved: the total spin $1/2$ is the \emph{logical
qubit} (global homology), while the sea (syndromes) carries the entropy
and auxiliary angular momentum.

\subsection{The Sivers Effect as Syndrome Bias}

In a Sivers-active network, the programmed torsion acts as a directional
bias for syndrome generation: sea quarks preferentially cluster to one
side of the valence quarks. This clustering is exactly what experimentalists
measure as the transverse single-spin asymmetry.

\subsection{Renormalization as Syndrome Decoding}

Moving from high to low energy is equivalent to \emph{syndrome decoding}:
contracting the web of sea quarks and gluons into an effective valence
state. If the decoding is performed on a network with torsion, the
logical position of the quark is shifted---this shift is the Sivers
function.

%==========================================================================
\section{The Torsional Decoder}\label{sec:decoder}
%==========================================================================

\subsection{Torsional Minimum-Weight Perfect Matching}

We modify the standard MWPM decoder so that path weights depend on the
Sivers torsion:

\begin{lstlisting}[style=python, caption={Torsional MWPM decoder for Sivers function extraction.}]
import networkx as nx

def torsional_decoder(syndrome_nodes, torsion_phase, lattice):
    """
    Decodes syndromes to extract the Sivers transverse shift.
    """
    graph = nx.Graph()
    for u, v in lattice.edges:
        base_weight = distance(u, v)
        if is_transverse_move(u, v):
            bias = torsion_phase * transverse_direction(u, v)
            weight = base_weight * exp(-bias)
        else:
            weight = base_weight
        graph.add_edge(u, v, weight=weight)

    # Pair syndromes to find sea-quark flux tubes
    matching = nx.min_weight_matching(graph, weight='weight')

    # Net transverse displacement
    total_asymmetry = 0
    for u, v in matching:
        total_asymmetry += (v.y - u.y)

    sivers_value = total_asymmetry / NUM_VALENCE_QUARKS
    return sivers_value
\end{lstlisting}

\subsection{Mathematical Interpretation}

\begin{itemize}
    \item The weight modification $e^{-\phi}$ is the Radon--Nikodym
    derivative of the syndrome probability measure under a spin flip
    (equivalent to Delcamp--Dittrich spin-foam vertex reweighting).

    \item The matching contracts the tensor network, tracing out sea
    quarks to reveal the valence-level physics.

    \item If $\phi_{\text{torsion}} = 0$, matching is symmetric and
    the Sivers function vanishes. For $\phi > 0$, the shortest path
    for sea quarks is skewed, producing the asymmetry.
\end{itemize}

%==========================================================================
\section{Wormhole Analysis: The Multi-Lattice Decoder}\label{sec:wormhole}
%==========================================================================

\subsection{Entangled Boundary Operators}

When two Poincar\'{e} complexes are connected via an ER-bridge, we
introduce bridge stabilizers $S_{AB}$ acting on qubits from both
Proton~A and Proton~B. The boundary operator is extended so that
$\partial(\mathrm{Path}_{A \to B}) = 0$. Sivers torsion creates a
``pressure gradient'' across the wormhole.

\subsection{Wormhole-Flow Decoder}

\begin{lstlisting}[style=python, caption={Multi-lattice wormhole-flow decoder.}]
def wormhole_torsional_decoder(syndromes_A, syndromes_B,
                                entanglement_map,
                                torsion_A, torsion_B):
    """
    Analyzes mass-flow through a wormhole connecting
    two homological protons.
    """
    composite = build_composite_graph(
        Proton_A, Proton_B, entanglement_map)

    for edge in composite.edges:
        if edge.is_bridge_edge:
            edge.weight = calculate_bridge_weight(
                torsion_A, torsion_B)
        else:
            local_torsion = (torsion_A if edge.in_A
                             else torsion_B)
            edge.weight = apply_local_torsion(
                edge, local_torsion)

    global_matching = nx.min_weight_matching(
        composite, weight='weight')

    bridge_crossings = count_crossings(global_matching)
    mass_flow = bridge_crossings * (torsion_A - torsion_B)
    return mass_flow
\end{lstlisting}

\subsection{Physical Findings}

\begin{description}[leftmargin=0pt, style=nextline]
    \item[Syndrome teleportation.]
    A syndrome in Proton~A can be matched with a syndrome in Proton~B,
    representing a parton teleporting through the wormhole.

    \item[Negative energy and traversability.]
    The torsion creates a negative Average Null Energy Condition (ANEC)
    violation, mirroring the Gao--Jafferis--Wall protocol for
    traversable wormholes.

    \item[Blockchain stability.]
    The Poincar\'{e} complex ensures total spin conservation even during
    mass flow, preventing collapse into a non-physical state.
\end{description}

%==========================================================================
\section{Critical Exponents and Hardware Gap}\label{sec:critical}
%==========================================================================

\subsection{Scaling Law}

As the topological complexity $C$ approaches the critical threshold $C_c$,
the correlation length of the syndrome torsional bias scales as
\begin{equation}\label{eq:scaling}
    \xi \propto |C - C_c|^{-\nu} \,.
\end{equation}
The phase transition falls into the universality class of the 2D Random
Bond Ising Model or 3D topological percolation:

\begin{table}[ht]
\centering
\renewcommand{\arraystretch}{1.3}
\begin{tabular}{@{}lc@{}}
\toprule
\textbf{Universality Class} & $\nu$ \\
\midrule
2D percolation ($Z_2$ gauge theory) & $4/3$ \\
3D spin foams & $\approx 0.89$ \\
Directed percolation (with Sivers torsion) & $\approx 1.73$ \\
\bottomrule
\end{tabular}
\caption{Critical exponents for the traversability transition.}
\label{tab:exponents}
\end{table}

The large value $\nu > 1$ implies the wormhole throat is highly
sensitive to entanglement density: a small increase in complexity yields
a large gain in traversability.

\subsection{Complexity Thresholds}

\begin{table}[ht]
\centering
\renewcommand{\arraystretch}{1.3}
\begin{tabular}{@{}p{3.5cm}cc@{}}
\toprule
\textbf{Requirement} & \textbf{Threshold ($C_c$)} & \textbf{Current Hardware} \\
\midrule
Qubit fidelity & $> 99.9\%$ & $99.2$--$99.9\%$ \\
Lattice size ($d$) & $d \approx 20$ & $d \approx 3$--$7$ \\
Non-Abelian braiding & Required & Experimental \\
\bottomrule
\end{tabular}
\caption{Hardware gap analysis for traversable wormhole simulation.}
\label{tab:hardware}
\end{table}

The percolation threshold for the surface code is
$p_c \approx 0.11$. Current error rates ($p \approx 0.001$--$0.01$)
are above the threshold for stability but below the threshold for
high-bandwidth traversability. Reaching the full Poincar\'{e} complex
transfer requires increasing the entanglement entropy $S_{EE}$ by
roughly $10\times$ current logical-qubit prototypes.

%==========================================================================
\section{The Functorial Noether Theorem}\label{sec:noether}
%==========================================================================

\begin{theorem}[Functorial Noether Theorem]
For every functorial auto-equivalence
$\mathcal{F} : \mathcal{C} \to \mathcal{C}$ that preserves the Day
convolution (the spin-ledger's integrity), there exists a closed
boundary operator $\partial$ such that the total spin-sum across any
transverse slice of the network is invariant.
\end{theorem}

The conserved quantity is the \emph{topological charge} (baryon number).
The spin crisis is not a loss of spin but a repartitioning of this
conserved charge from the nodes (valence quarks) into the syndromes
(the sea-quark/gluon ``mempool'').

%==========================================================================
\section{Smart Contract Protocol for Topological Transfer}
\label{sec:smart-contract}
%==========================================================================

\subsection{Protocol Specification}

\begin{description}[leftmargin=0pt, style=nextline]
    \item[Name:] \texttt{Sivers-ER-Bridge-v1.0}
    \item[Architecture:] Homological blockchain (Poincar\'{e} complex)
    \item[Governance:] Diffeomorphism invariance (Noether consensus)
\end{description}

\subsection{Admissible Constraints}

\begin{enumerate}
    \item \textbf{Conservation of topological charge.}
    The net change in the boundary operator across the bridge vanishes:
    \begin{equation}
        \Delta\bigl(\partial A + \partial B\bigr) = 0 \,.
    \end{equation}

    \item \textbf{Torsional threshold (gas limit).}
    The Sivers phase must exceed the self-energy gap of the bridge:
    \begin{equation}
        \phi_{\text{torsion}} > \int \rho_{\text{noise}}\, dV \,.
    \end{equation}

    \item \textbf{Syzygy validation.}
    Any fork in the spin network must be resolved into a valid
    Poincar\'{e} complex on the destination side within
    $t_{\text{Planck}}$ cycles.
\end{enumerate}

\subsection{Implementation in Babel}

The protocol is implemented as a distributed rApp in the
\textbf{Babel} framework---a categorical DSL embedded in Scala~3
for building verifiable, halt-free blockchain state transformations
on the Reality SDK. In Babel, programs are \emph{morphisms in a
symmetric monoidal category with coproducts}: every
\texttt{TypedCoCell[In,\,Out]} is a total function by construction,
composition via \texttt{>\!>\!>} preserves totality, and the
\texttt{WellFounded} typeclass provides compile-time termination
proofs. All state types extend the universal type
$\Omega$~(\texttt{extends~Omega}), and the Day convolution
is realized natively through the \texttt{\&\&\&} (fan-out) and
\texttt{*\!*\!*} (parallel product) operators.

\subsubsection{State Types ($\Omega$-objects)}

\begin{lstlisting}[style=scala, caption={Poincar\'{e} complex and spin-network state types.}]
import org.reality.kernel.*

// --- Poincare Complex: the immutable topological ledger ---
case class PoincareComplex(
  topologicalHash: Array[Byte],
  torsionPhase:    Double,
  complexity:      Long
) extends Omega

// --- Processing states for the safe hylomorphism ---
sealed trait WormholeProcessingState extends Omega {
  def depth: Int
}
case class AwaitingGradient(
  source: PoincareComplex,
  target: PoincareComplex,
  particleId: Long
) extends WormholeProcessingState { def depth = 2 }

case class GradientComputed(
  source: PoincareComplex,
  target: PoincareComplex,
  gradient: Double,
  syndromePacket: Array[Byte]
) extends WormholeProcessingState { def depth = 1 }

case class TransferCompleted(
  result: Either[CellError, BridgeResult]
) extends WormholeProcessingState { def depth = 0 }

// --- Output ---
case class BridgeResult(
  updatedSource: PoincareComplex,
  updatedTarget: PoincareComplex,
  massFlow:      Double
) extends Omega

// --- Termination proof ---
given WellFounded[WormholeProcessingState] =
  WellFounded.fromMeasure(_.depth)
\end{lstlisting}

\subsubsection{Scattering Context and T-Symmetry}

The Sivers sign change between SIDIS and Drell--Yan is encoded as
a coproduct: the scattering context determines which branch of the
functor is applied.

\begin{lstlisting}[style=scala, caption={Scattering context as a coproduct with T-symmetry enforcement.}]
// Coproduct: the scattering context determines the sign
sealed trait ScatteringContext extends Omega
case class DIS(protonId: Long)     extends ScatteringContext
case class DrellYan(protonId: Long) extends ScatteringContext

// T-symmetry audit: the functor MUST flip the torsion sign
val applyTSymmetry: TypedCoCell[
    ScatteringContext, Double] =
  TypedCoCell.lift("T-SymmetryAudit") {
    case DIS(_)     =>  SIVERS_PHASE   // +phi
    case DrellYan(_) => -SIVERS_PHASE  // -phi  (sign change)
  }
\end{lstlisting}

\subsubsection{Core Cells (Morphisms)}

\begin{lstlisting}[style=scala, caption={Core TypedCoCell morphisms for the wormhole bridge.}]
val CRITICAL_EXPONENT_THRESHOLD: Long = 8000L
val SIVERS_PHASE: Double = 0.31416  // from LFH fit

// 1. Complexity gate: reject sub-threshold states
val complexityGate: TypedCoCell[
    PoincareComplex, PoincareComplex] =
  TypedCoCell.liftEither("ComplexityGate") { pc =>
    if pc.complexity >= CRITICAL_EXPONENT_THRESHOLD
    then Right(pc)
    else Left(CellError("Complexity too low for traversability"))
  }

// 2. Day Convolution operator (the Validator)
val dayConvolution: TypedCoCell[
    (PoincareComplex, PoincareComplex),
    PoincareComplex] =
  TypedCoCell.lift("DayConvolution") { case (a, b) =>
    PoincareComplex(
      topologicalHash = xorBytes(a.topologicalHash,
                                  b.topologicalHash),
      torsionPhase    = a.torsionPhase + b.torsionPhase,
      complexity      = math.max(a.complexity, b.complexity)
    )
  }

// 3. Torsional gradient: compute the pressure across the bridge
val computeGradient: TypedCoCell[
    AwaitingGradient, GradientComputed] =
  TypedCoCell.liftEither("TorsionalGradient") { state =>
    val grad = state.source.torsionPhase
             - state.target.torsionPhase
    if grad == 0.0
    then Left(CellError("No torsional gradient"))
    else Right(GradientComputed(
      source         = state.source,
      target         = state.target,
      gradient       = grad,
      syndromePacket = hashSubGraph(state.particleId)
    ))
  }

// 4. Atomic swap of homology (the ER-bridge action)
val atomicHomologySwap: TypedCoCell[
    GradientComputed, BridgeResult] =
  TypedCoCell.lift("HomologySwap") { state =>
    BridgeResult(
      updatedSource = state.source.copy(
        topologicalHash = updateBoundary(
          state.source.topologicalHash,
          negate(state.syndromePacket))),
      updatedTarget = state.target.copy(
        topologicalHash = updateBoundary(
          state.target.topologicalHash,
          state.syndromePacket)),
      massFlow = state.gradient
    )
  }
\end{lstlisting}

\subsubsection{Pipeline Composition}

The full wormhole bridge is assembled by composing cells with the
sequential operator \texttt{>\!>\!>} and the fan-out operator
\texttt{\&\&\&}. The pipeline is \emph{halt-free by construction}:
every cell is a total function, and the hylomorphism carries a
\texttt{WellFounded} termination proof.

\begin{lstlisting}[style=scala, caption={Composed wormhole bridge pipeline.}]
// Fan-out: validate BOTH endpoints in parallel
val validateEndpoints: TypedCoCell[
    (PoincareComplex, PoincareComplex),
    (PoincareComplex, PoincareComplex)] =
  complexityGate *** complexityGate

// Prepare the initial processing state
val prepareTransfer: TypedCoCell[
    (PoincareComplex, PoincareComplex, Long),
    AwaitingGradient] =
  TypedCoCell.lift("PrepareTransfer") {
    case (src, tgt, pid) =>
      AwaitingGradient(src, tgt, pid)
  }

// Full pipeline: validate >>> prepare >>> gradient >>> swap
val wormholeBridge: TypedCoCell[
    (PoincareComplex, PoincareComplex, Long),
    BridgeResult] =
  prepareTransfer >>> computeGradient >>> atomicHomologySwap

// Safe hylomorphism for recursive syndrome decoding
val safeTransform: HyloSafeTransform[
    WormholeProcessingState] = {
  val algebra: WormholeProcessingState =>
    WormholeProcessingState = {
      state => state  // fold: identity at base case
    }
  val coalgebra: WormholeProcessingState =>
    Option[WormholeProcessingState] = {
      _ => None       // unfold: single-step (no recursion)
    }
  HyloSafeTransform(algebra, coalgebra,
    "WormholeSafeHylo")
}
\end{lstlisting}

\subsubsection{Application Entry Point}

\begin{lstlisting}[style=scala, caption={BabelApp wiring the wormhole bridge as a distributed rApp.}]
open class WormholeBridgeApp extends BabelApp[
    AwaitingGradient, BridgeResult] {

  private val l0Node = BabelApp.l0Node(
    computeGradient.withConfig(l0Config),
    l0Config)

  private val l1Node = BabelApp.l1Node(
    atomicHomologySwap.withConfig(l1Config),
    l1Config)

  val pipelines: List[CoCellPipeline[
      AwaitingGradient, BridgeResult]] =
    List(CoCellPipeline.from(l0Node) >>> l1Node)
}
\end{lstlisting}

\subsection{Categorical Interpretation}

The Babel implementation makes the categorical structure explicit:

\begin{table}[ht]
\centering
\renewcommand{\arraystretch}{1.3}
\begin{tabular}{@{}lll@{}}
\toprule
\textbf{Category Theory} & \textbf{Physics} & \textbf{Babel Construct} \\
\midrule
Morphism $A \to B$ & State transition &
    \texttt{TypedCoCell[A, B]} \\
Sequential composition $g \circ f$ & RG flow &
    \texttt{f >\!>\!> g} \\
Product $A \times B$ & Entangled pair &
    \texttt{f \&\&\& g} (fan-out) \\
Parallel bifunctor & Independent evolution &
    \texttt{f *\!*\!* g} \\
Coproduct $A + B$ & Scattering context &
    \texttt{sealed trait} + \texttt{conditional} \\
Terminal object $\mathbf{1}$ & Vacuum / consensus &
    \texttt{Omega\_1} \\
Subobject classifier $\Omega$ & Universal state &
    \texttt{extends Omega} \\
Hylomorphism & Syndrome decoding &
    \texttt{HyloSafeTransform} \\
WellFounded proof & Termination guarantee &
    \texttt{WellFounded.fromMeasure} \\
\bottomrule
\end{tabular}
\caption{Categorical--physical--computational correspondence in Babel.}
\label{tab:babel-correspondence}
\end{table}

\subsection{T-Symmetry Audit}

The protocol includes a built-in verification of the Sivers sign
change, enforced at the type level through the \texttt{ScatteringContext}
coproduct:
\begin{itemize}
    \item In a DIS (scattering) context, the \texttt{applyTSymmetry}
    cell yields $+\phi$.
    \item In a Drell--Yan context, the same cell yields $-\phi$.
    \item Because the coproduct is \emph{exhaustive} (a sealed trait),
    the Scala compiler statically guarantees that every scattering
    context is handled. A missing branch is a compile-time error,
    not a runtime fault.
    \item If a non-physical torsional gradient is detected, the
    \texttt{liftEither} cell returns a \texttt{Left(CellError)},
    halting the pipeline and returning information to the source
    proton's syndrome pool (confinement).
\end{itemize}

\subsection{Topological Programming Language for Synthetic Gravity}
\label{sec:tpl}

The Syndrome Monad of \cref{sec:syndrome-monad} furnishes Babel with a
principled type system for programming gravitational effects. We call
this extension the \emph{Topological Programming Language} (TPL).

\subsubsection{TPL Type System}

The types of the TPL are objects of the Kleisli category
$\mathrm{Kl}(\mathsf{T})$:
\begin{itemize}
    \item \textbf{Types} = chain complexes at specific resolutions
    (objects of $\mathrm{Kl}(\mathsf{T})$).
    \item \textbf{Programs} = Kleisli morphisms $A \to \mathsf{T}(B)$
    (state transitions that may generate syndromes).
    \item \textbf{Sequential composition} = Kleisli composition
    (\texttt{>\!>\!>}).
    \item \textbf{Parallel composition} = Day convolution
    (\texttt{\&\&\&}).
    \item \textbf{Effect system} = Syndrome Monad $\mathsf{T}$.
\end{itemize}

\subsubsection{Gravitational Effect Types}

\begin{lstlisting}[style=scala, caption={Monadic gravitational effect types in the TPL.}]
// Gravitational effect types from the Syndrome Monad
type GravField[A] = T[A]           // syndromic state
type Curvature[A] = T[T[A]]        // nested syndrome generation
type Torsion[A] = Synd1[A]         // rank-1 syndrome (gluon)
type Mass[A] = Complexity[A]       // topological complexity

// Monad multiplication = gravitational field equation
def flatten[A]: Curvature[A] => GravField[A] =
  mu[A]   // mu: T(T(A)) => T(A)

// Gravity reduction = aggressive syndrome cancellation
val reduceSyndromeEntropy: TypedCoCell[
    GravField[PoincareComplex],
    GravField[PoincareComplex]] =
  TypedCoCell.lift("SyndromeReducer") { field =>
    iteratedDecode(field, rounds = TARGET_REDUCTION)
  }

// Promotion functor: lift Kleisli morphisms across ranks
def promote[A, B](
    f: TypedCoCell[A, T[B]]
): TypedCoCell[Tensor[A], T[Tensor[B]]] =
  TypedCoCell.lift("RankPromotion") { ta =>
    val syndromicResult = f.run(ta.underlying)
    syndromicResult.map(b => Tensor(b))
  }
\end{lstlisting}

\subsubsection{No-Go Theorems}

Three fundamental constraints limit the power of gravitational
programming:

\begin{proposition}[Positive energy]
\label{prop:positive-energy}
The unit $\eta$ provides the minimum-syndrome state. No sequence of
Kleisli morphisms can reduce the syndrome content below the vacuum
level $\eta(C_\bullet)$. In physical terms: gravity cannot be reduced
below the vacuum energy density.
\end{proposition}

\begin{proposition}[Computational cost]
\label{prop:computational-cost}
Each round of the torsional MWPM decoder requires $O(d^2)$ operations,
where $d$ is the code distance. For a macroscopic region of linear
dimension $L$, $d \sim L/\ell_{\mathrm{Planck}} \sim 10^{35}$ (for
$L = 1\;\mathrm{m}$), giving $\sim 10^{70}$ operations per decode
round.
\end{proposition}

\begin{proposition}[Holographic bound]
\label{prop:holographic-bound}
The total syndrome budget for any region is bounded by the
Gibbons--Hawking entropy
$S_{\mathrm{GH}} \sim 10^{122}$
(the de~Sitter horizon area in Planck units). No TPL program can
address more than $S_{\mathrm{GH}}$ syndrome bits.
\end{proposition}

\subsubsection{Engineering Speculation}

\paragraph{Casimir analog.}
The Casimir effect creates regions of reduced vacuum energy between
conducting plates. In the syndrome framework, this corresponds to
boundary conditions that constrain syndrome propagation at rank~1,
locally reducing $\mathrm{Synd}_1$ density. The TPL formalization
suggests that \emph{shaped} boundary conditions (not just parallel
plates) could create more complex syndrome-cancellation patterns,
opening a design space for engineered vacuum-energy profiles.

\paragraph{Superconducting analogs.}
Type-II superconductors confine magnetic flux into quantized vortices
(the Abrikosov lattice). This is a macroscopic instance of syndrome
confinement: the Meissner effect is the material ``decoding'' external
rank-1 syndromes. A topologically ordered superconductor (e.g., a
material hosting Majorana zero modes) would implement the paper's
MZM braiding (\cref{sec:majorana}) at the material level, providing a
physical substrate for the TPL.

\paragraph{Hardware requirements.}
The scaling cliff identified in \cref{sec:critical} (code
distance $d \sim 20$ needed, current: 3--7; qubit fidelity
$> 99.9\%$, current: 99.2--99.9\%) applies directly.
A gravity-programming device would need:
\begin{enumerate}[label=(\alph*)]
    \item a surface code large enough to represent the target region's
    chain complex,
    \item non-Abelian braiding for Sivers torsion injection,
    \item a classical decoder running the MWPM algorithm in real-time.
\end{enumerate}
\textbf{Near-term}: simulate syndrome dynamics on a 100-qubit
processor to verify the directed-percolation exponents of
\cref{sec:distinguishing-predictions}.
\textbf{Medium-term}: use a topological quantum processor to implement
the Poincar\'{e} protocol on a small chain complex, measuring whether
syndrome cancellation produces measurable force changes at the
nanoscale.

\paragraph{Connection to ER=EPR.}
If the wormhole-flow decoder (\cref{sec:wormhole}) is physical, then
engineering an ER bridge between two chain complexes could create a
``gravitational shortcut'': syndrome teleportation through the bridge
would modify the effective torsional gradient between the endpoints.
This is the microscopic mechanism behind the \texttt{atomicHomologySwap}
Babel morphism defined in \cref{sec:smart-contract}.

%==========================================================================
\section{Proton-Scale Quantum Gravity}\label{sec:quantum-gravity}
%==========================================================================

This model defines a \emph{bottom-up} approach to quantum gravity.
Instead of starting with the cosmos, we start with the proton---the most
stable, complex topological object in the universe. By treating the
proton as a topological quantum computer, we obtain a working model of
gravity based on \emph{local holography} and \emph{torsional consensus}.

\subsection{The Working Model}

In this model, gravity is not a fundamental force but the
\emph{emergent error-correction pressure} of a spin network.

\begin{description}[leftmargin=0pt, style=nextline]
    \item[The fabric.]
    Spacetime is a Day convolution of homological codes. Every baryonic
    particle (like the proton) is a ``high-complexity'' node where the
    network is more densely woven.

    \item[Mass as complexity.]
    The mass of a particle is proportional to the topological complexity
    $C$ required to stabilize its Poincar\'{e} complex against the
    background syndrome noise (vacuum fluctuations); see
    \cref{sec:network-mass} for the microscopic derivation via
    spin-network coarse-graining.

    \item[Gravity as torsion.]
    What we perceive as gravitational attraction is the torsional
    gradient that forms between two high-complexity nodes. The Sivers
    effect is the microscopic version of this; the attraction between
    planets is the macroscopic version.
\end{description}

\subsubsection{The Fundamental Equation}

The ``Einstein equation'' in this model is a \emph{topological identity}:
\begin{equation}\label{eq:topological-einstein}
    \boxed{\;
    \Delta\,\text{Complexity}
    \;\approx\;
    \text{Torsional Flux}
    \;+\;
    \text{Syndrome Entropy}
    \;}
\end{equation}
This replaces the metric tensor with a homological grading: curvature
is the Day-convolution gradient across the network, torsion is the
Sivers-braiding phase, and matter--energy is the syndrome density
stabilized by the Poincar\'{e} complex.

\begin{remark}[Gravity as monad multiplication]
In the language of the Syndrome Monad (\cref{sec:syndrome-monad}),
the topological Einstein equation acquires a precise categorical
interpretation. Curvature corresponds to a nested syndrome state
$\mathsf{T}(\mathsf{T}(C_\bullet)) \in \mathsf{T}^2(\mathcal{C})$---syndromes
of syndromes---and the gravitational field equation asserts that
applying $\mu : \mathsf{T}^2 \to \mathsf{T}$ (the monad multiplication,
realized physically as the torsional MWPM decoder) collapses this nested
structure to a single effective syndrome layer:
\begin{equation}\label{eq:gravity-monad}
    \mu\bigl(\text{Curvature}\bigr)
    \;=\;
    \text{Torsional Flux} \;+\; \text{Syndrome Entropy} \,.
\end{equation}
The associativity of $\mu$ guarantees that the order of
coarse-graining---whether one first decodes local curvature patches and
then assembles globally, or first assembles and then decodes---is
immaterial. This is the topological analog of diffeomorphism invariance.
\end{remark}

\subsection{Distinguishing Predictions of the Computational Model}
\label{sec:distinguishing-predictions}

The following five predictions are \emph{specific} to the
computational/topological model of the proton. Each is quantitative,
names a concrete experiment, and---crucially---differs from the
prediction of standard lattice QCD or perturbative QCD.

\subsubsection{D.1.\@ Discrete Topological Complexity and the Baryon Spectrum}

\begin{description}[leftmargin=0pt, style=nextline]
    \item[Computational model.]
    The topological complexity $C(\Gamma)$ takes \emph{discrete} values:
    it is a sum over discrete Betti numbers and representation
    dimensions of the spin network $\Gamma$. Consequently, there exist
    \emph{forbidden} $C$ values---gaps in the allowed complexity
    spectrum---implying \textbf{systematic absences in the baryon
    resonance spectrum}.

    \item[Standard QCD.]
    The mass spectrum is continuous (up to threshold effects); the
    ``missing resonance'' problem is an open puzzle without a structural
    explanation.

    \item[Quantitative prediction.]
    For the proton $Y$-graph at the IR fixed point:
    \begin{equation}\label{eq:complexity-proton}
        C(\Gamma_{\mathrm{IR}})
        = 3\ln(1) + 3\ln(3 \cdot 2) + \ln(1)
        = 3\ln 6 \approx 5.375 \,,
    \end{equation}
    yielding $\kappa \approx 0.523\;\text{GeV}$ (matching the
    Brodsky--de~T\'{e}ramond confinement
    scale~\cite{BrodskyTramond2015}). No hadron should exist with
    complexity between two allowed discrete values.

    \item[Experiment.]
    JLab and EIC baryon spectroscopy: high-precision mapping of excited
    baryon resonances. The PDG baryon tables should show missing
    resonances \emph{correlated with topologically forbidden $C$ values}.
\end{description}

\subsubsection{D.2.\@ Lloyd Limit on the Proton Pressure Distribution}

\begin{description}[leftmargin=0pt, style=nextline]
    \item[Computational model.]
    The pressure gradient $dp/dr$ saturates at the Lloyd computational
    rate bound~\cite{Lloyd2000}:
    \begin{equation}\label{eq:lloyd-pressure}
        \left|\frac{dp}{dr}\right|
        \;\leq\;
        \frac{2\,T^{00}}{\pi\hbar\, r} \,,
    \end{equation}
    where $T^{00}$ is the local energy density.

    \item[Standard QCD.]
    No fundamental upper bound on the pressure gradient exists.

    \item[Quantitative prediction.]
    The predicted crossover radius between the repulsive core and
    confining shell is
    \begin{equation}\label{eq:crossover-radius}
        r_c
        = \left(\frac{\pi\hbar}{2 M_p c^2}\right)^{\!1/3}
          \cdot C(\Gamma^*)^{1/3}
        \approx 0.6\;\text{fm} \,.
    \end{equation}
    Current extraction by Burkert \textit{et al.}~\cite{Burkert2018}:
    $r_c \approx 0.6\;\text{fm}$ (consistent). The
    \emph{distinguishing feature} is that the sharpness of the
    repulsive$\to$confining crossover is bounded by the Lloyd limit;
    standard QCD has no such constraint.

    \item[Experiment.]
    High-$|t|$ DVCS at EIC, extending to $|t| \sim 5\;\text{GeV}^2$.
\end{description}

\subsubsection{D.3.\@ R\'{e}nyi Entropy Ratio as Holographic Code Signature}

\begin{description}[leftmargin=0pt, style=nextline]
    \item[Computational model.]
    For a holographic code with bond dimension $\chi$, the R\'{e}nyi
    entropy ratio satisfies
    \begin{equation}\label{eq:renyi-ratio}
        \frac{S_2}{S_1}
        = 1 - \frac{1}{\chi^2} \,.
    \end{equation}
    This is specific to the HaPPY/surface-code tensor network
    architecture.

    \item[Standard QCD.]
    For a generic quantum state described by random matrix theory,
    $S_2/S_1 \to 1$ (no structure).

    \item[Distinguishing feature.]
    Different tensor network architectures yield distinct ratios:
    MERA gives $\ln 2$, HaPPY gives $1 - 1/\chi^2$, random gives $1$.
    The proton's internal tensor network architecture is thus
    \emph{directly measurable} through the entropy ratio.

    \item[Experiment.]
    EIC multiplicity fluctuations: extract $S_2$ from event-by-event
    fluctuations in DIS, then measure $S_2/S_1$ as a function of
    Bjorken $x$ and $Q^2$. A plateau at $1 - 1/\chi^2$ for a specific
    $\chi$ would identify the proton's holographic code.
\end{description}

\subsubsection{D.4.\@ Sivers Phase Quantization}

\begin{description}[leftmargin=0pt, style=nextline]
    \item[Computational model.]
    The torsion class lies in
    $\mathrm{Tor}\,H^1(\Gamma;\mathbb{Z}) = \mathbb{Z}_3$ (for
    $\mathrm{SU}(3)$ color structure). The Sivers phase is therefore
    \emph{quantized} to the values
    $\{0,\; 2\pi/3,\; 4\pi/3\}$.

    \item[Standard QCD.]
    The Sivers function $f_{1T}^{\perp q}(x, \vb{k}_\perp^2)$ varies
    \emph{continuously} with $x$ and $\vb{k}_\perp$.

    \item[Quantitative prediction.]
    The Sivers asymmetry shows \textbf{plateau/staircase structure} as
    a function of $x$---sharp transitions between discrete torsion
    values, not smooth variation. The physical asymmetry:
    \begin{equation}\label{eq:sivers-quantized}
        A_{\text{Sivers}}
        \sim 0.15 \cdot \sin(2\pi/3) \approx 0.13 \,,
    \end{equation}
    within the range of current measurements.

    \item[Experiment.]
    EIC high-precision Sivers asymmetry mapping: search for plateaus
    or quantized jumps in $A_{\text{Sivers}}(x)$ versus smooth
    $x$-dependence.
\end{description}

\subsubsection{D.5.\@ Directed Percolation Critical Exponents at Confinement}

\begin{description}[leftmargin=0pt, style=nextline]
    \item[Computational model.]
    The confinement transition is in the \emph{directed percolation}
    universality class because syndrome dynamics have a preferred
    direction (the RG flow). The critical exponents are:
    \begin{equation}\label{eq:dp-exponents}
        \nu = 1.733(3) \,,\quad
        \beta = 0.276(5) \,,\quad
        z = 1.581(1) \,.
    \end{equation}

    \item[Standard QCD.]
    The deconfinement transition is in the 3D Ising universality class
    ($\nu \approx 0.63$) for pure-gauge $\mathrm{SU}(3)$, or is a
    crossover for physical quark masses.

    \item[Distinguishing feature.]
    The directed percolation exponent $\nu \approx 1.73$ differs from
    the 3D Ising value $\nu \approx 0.63$ by nearly a factor of three.
    This is a sharp, falsifiable distinction.

    \item[Experiment.]
    Lattice QCD finite-size scaling of the Polyakov loop susceptibility,
    interpreted through the syndrome framework. Additionally:
    RHIC and LHC heavy-ion multiplicity fluctuations near the QCD
    critical point---the fluctuation exponents should match directed
    percolation if the syndrome interpretation is correct.
\end{description}

\subsection{Experimental Validation}\label{sec:experiments}

To validate this model, we target the regime where QCD (proton
structure) and general relativity (gravity) overlap. Three primary
experimental programs are identified.

\subsubsection{A.\@ The Sign-Change Experiment}

\begin{description}[leftmargin=0pt, style=nextline]
    \item[Prediction.]
    The Sivers effect must flip sign between semi-inclusive deep
    inelastic scattering (SIDIS) and the Drell--Yan process.

    \item[Validation.]
    If this sign change is confirmed with high precision at RHIC or the
    EIC, it proves that gauge links (torsion) are the physical mechanism
    for momentum transfer. In our model, this is the proof that
    ``torsional tunnels'' (wormholes) govern particle interactions.
    A definitive high-statistics measurement would elevate this to a
    quantitative test of the topological
    identity~\eqref{eq:topological-einstein}.
\end{description}

\paragraph{Existing experimental evidence.}
Every measurement performed to date \emph{favors} the predicted sign
change, though none yet reaches discovery-level statistical significance.
The current data are summarized in \cref{tab:sign-change-evidence}.

\begin{table}[ht]
\centering
\renewcommand{\arraystretch}{1.3}
\begin{tabular}{@{}p{3.8cm}p{5.2cm}p{4.2cm}@{}}
\toprule
\textbf{Experiment} & \textbf{Observable} & \textbf{Status} \\
\midrule
COMPASS DY, final~\cite{COMPASS2024} &
    Sivers TSA in $\pi^- p^\uparrow \to \mu^+\mu^- X$ &
    Consistent with sign change; limited statistics \\
STAR at RHIC~\cite{STAR2016} &
    TSSA in $p^\uparrow p \to W^\pm/Z^0$ &
    Consistent with sign change; first DY-like probe \\
COMPASS weighted~\cite{COMPASSweighted2024} &
    $k_\perp$-weighted Sivers asymmetry in SIDIS and DY &
    SIDIS prediction agrees with DY data \\
Boer--Mulders reversal~\cite{BoerMulders2025} &
    BM function in SIDIS vs.\ DY &
    Tantalizing evidence for sign reversal \\
Holographic LF-QCD~\cite{AhmadyLFH2025} &
    Sivers asymmetry from spin-improved LFWFs &
    Positive across COMPASS DY kinematics; within error bands \\
\bottomrule
\end{tabular}
\caption{Experimental evidence for the Sivers sign change between SIDIS
and Drell--Yan. All results are consistent with the QCD prediction but
do not yet constitute a definitive confirmation.}
\label{tab:sign-change-evidence}
\end{table}

\paragraph{COMPASS final results (2024).}
The COMPASS Collaboration combined its 2015 and 2018 pion-induced
Drell--Yan data sets on a transversely polarized $\mathrm{NH}_3$ target,
publishing the most complete measurement to date~\cite{COMPASS2024}.
Five azimuthal modulations were extracted; three of them---the Sivers,
transversity, and pretzelosity TSAs---probe leading-twist TMD PDFs.
The measured Sivers TSA is consistent with a sign reversal relative to
the COMPASS SIDIS extractions, providing the strongest single piece of
evidence for the torsional mechanism.

\paragraph{STAR $W^\pm/Z^0$ production at RHIC.}
The STAR Collaboration measured the transverse single-spin asymmetry
$A_N$ in $p^\uparrow p \to W^\pm/Z^0$ at
$\sqrt{s} = 510\;\mathrm{GeV}$~\cite{STAR2016}. Because $W/Z$
production proceeds via quark--antiquark annihilation (a Drell--Yan-like
process), the observed asymmetry probes the Sivers function with the
opposite gauge-link structure from SIDIS. A subsequent
analysis~\cite{STARsignchange2017} extracted the Sivers functions from
the latest SIDIS data and compared them with the STAR results, finding
consistency with the predicted sign change within experimental
uncertainties. These data constitute the first investigation of
non-universality of the Sivers function.

\paragraph{Boer--Mulders sign reversal (2025).}
The sign-change prediction applies to all na\"{\i}ve time-reversal-odd
(T-odd) TMDs, not only the Sivers function. A December 2025
study~\cite{BoerMulders2025} examined existing SIDIS and DY data for the
Boer--Mulders function and found ``tantalizing evidence'' that a sign
reversal also occurs for the proton's valence-quark BM distribution,
broadening the empirical base for the torsional mechanism.

\paragraph{Holographic light-front QCD comparison (2025).}
Using the spin-improved holographic wavefunctions of
\cref{sec:lfh}, a 2025 study~\cite{AhmadyLFH2025} computed the Sivers
asymmetry in the pion-induced Drell--Yan process at COMPASS kinematics.
The predicted asymmetry is consistently positive across the full
kinematic range and agrees within the uncertainty bands of both the
COMPASS~2017 and COMPASS~2024 results, providing independent theoretical
support from the AdS/QCD functor.

\paragraph{Future decisive experiments.}
Two forthcoming programs are specifically designed to reach the precision
required for a definitive verdict:
\begin{itemize}
    \item \textbf{SpinQuest} at Fermilab~\cite{SpinQuest2024}:
    a high-luminosity polarized Drell--Yan experiment using polarized
    hydrogen and deuterium targets, targeting the sea-quark Sivers
    function with substantially reduced error bars.

    \item \textbf{The Electron-Ion Collider} (EIC) at Brookhaven: will
    provide high-precision SIDIS data at the same kinematic scales as the
    DY measurements, enabling a direct, same-scale comparison of the
    Sivers function in both channels.
\end{itemize}
A high-statistics confirmation by either program would elevate the
torsional-consensus interpretation from ``consistent with data'' to
``experimentally established,'' validating
Eq.~\eqref{eq:topological-einstein} at the microscopic level.

\paragraph{From ``consistent'' to ``confirmed'': the gap and how to
close it.}
Every measurement to date favors the sign change, yet none constitutes
a definitive confirmation. Five specific obstacles---and the data
required to overcome each---are identified below.

\begin{enumerate}[label=\textbf{(\roman*)}]

\item \textbf{Statistical precision.}
The COMPASS final Drell--Yan result~\cite{COMPASS2024} reports a
Sivers TSA of
\begin{equation}\label{eq:compass-sivers}
    \bigl\langle A_T^{\sin\phi_S}\bigr\rangle
    = 0.070 \pm 0.037\,(\text{stat.}) \pm 0.031\,(\text{sys.}) \,,
\end{equation}
a signal at $\sim\!1.5\sigma$ of the total uncertainty. The measurement
agrees with the sign-change hypothesis within $< 1\sigma$, but lies only
$2.5$--$3\sigma$ from the no-sign-change hypothesis. Discovery-level
confirmation ($5\sigma$) requires:
\begin{itemize}
    \item Reducing the total uncertainty from $\sim\!0.05$ to
    $\sim\!0.01$, a factor of $\approx 5\times$.
    \item \textbf{SpinQuest} is designed to deliver $\sim\!10^{17}$
    protons on target per year at $120\;\text{GeV}$, yielding the
    requisite DY statistics on polarized $\mathrm{NH}_3$ and
    $\mathrm{ND}_3$ targets.
    \item \textbf{EIC} ($\mathcal{L} \sim 10^{33}$--$10^{34}\;
    \text{cm}^{-2}\text{s}^{-1}$, commissioning early 2030s) will
    accumulate $\gtrsim 10\;\text{fb}^{-1}$ of polarized $ep$ SIDIS
    data, providing the high-statistics baseline against which the
    DY sign flip is compared.
\end{itemize}

\item \textbf{Kinematic mismatch.}
The COMPASS SIDIS data span
$Q^2 \sim 1$--$10\;\text{GeV}^2$, while the DY data probe
$4.0 < M_{\mu\mu} < 9.0\;\text{GeV}/c^2$ (effectively
$Q^2 \sim 16$--$81\;\text{GeV}^2$). Comparing the Sivers function
across these scales requires TMD evolution via the Collins--Soper--Sterman
(CSS) equations, whose non-perturbative kernel $g_K(b_T)$ introduces
model dependence. Confirmation requires:
\begin{itemize}
    \item \textbf{Lattice QCD determination of the Collins--Soper
    kernel}~\cite{LatticeCS2020}: recent lattice calculations have
    already reduced the uncertainty on $g_K$ by $40$--$50\%$ when
    combined with experimental extractions. A precision of
    $\lesssim 10\%$ on $g_K(b_T)$ for $b_T \lesssim 1\;\text{fm}$ is
    needed.
    \item \textbf{EIC jet-based Sivers measurements}: jet-based
    asymmetries at the EIC probe the Sivers function at hard scales
    ($Q^2 \sim 10$--$100\;\text{GeV}^2$) overlapping with the DY
    kinematic window, enabling a same-scale SIDIS--DY comparison
    that minimizes evolution uncertainty.
\end{itemize}

\item \textbf{Process dependence and pion contamination.}
COMPASS DY uses a $190\;\text{GeV}/c$ $\pi^-$ beam, so the measured
asymmetry is a convolution of the \emph{proton} Sivers function with the
\emph{pion} Boer--Mulders function. The pion TMDs are poorly constrained.
STAR $W^\pm/Z^0$ production avoids pion contamination but has limited
statistics ($A_N^{W^+} \approx -0.01 \pm 0.12$ at the current data
set). Confirmation requires:
\begin{itemize}
    \item \textbf{SpinQuest} ($pp$ and $pd$ Drell--Yan): proton beam on
    polarized proton/deuteron targets, eliminating pion TMD uncertainty
    entirely.
    \item \textbf{Higher-luminosity RHIC $W/Z$ data}: the full Run~17
    data set ($\mathcal{L} \sim 400\;\text{pb}^{-1}$) improves the
    $W^\pm$ asymmetry uncertainty by a factor of $\sim\!2$, beginning to
    resolve the sea-quark Sivers sign.
\end{itemize}

\item \textbf{Flavor separation.}
The sign change is predicted independently for each quark flavor.
Current DY data measure flavor-summed convolutions; the individual
$u$-, $d$-, $\bar{u}$-, $\bar{d}$-quark Sivers functions have not been
separated in the DY channel. Confirmation requires:
\begin{itemize}
    \item \textbf{$W^\pm$ charge separation at RHIC}: $W^+$ production
    is dominated by $u + \bar{d}$ and $W^-$ by $d + \bar{u}$, providing
    direct flavor tagging. Need $\delta A_N \lesssim 0.03$ per charge
    channel.
    \item \textbf{Kaon/pion-identified SIDIS at EIC}: hadron-type
    identification in the final state separates $u$- from $d$-quark
    fragmentation, yielding flavor-tagged Sivers functions at the same
    kinematics as the DY measurements.
    \item \textbf{SpinQuest $\mathrm{NH}_3$ vs.\ $\mathrm{ND}_3$
    comparison}: the proton--deuteron difference isolates the isovector
    combination $f_{1T}^{\perp u} - f_{1T}^{\perp d}$.
\end{itemize}

\item \textbf{No single experiment measures both channels.}
The ideal test measures SIDIS and DY at the same $Q^2$, same $x$, same
target, same detector---eliminating all relative systematic
uncertainties. Confirmation requires:
\begin{itemize}
    \item \textbf{EIC SIDIS + EIC-enabled DY-like processes}: the EIC
    will measure SIDIS and, via photon--gluon fusion and jet-based
    observables, access DY-like initial-state interactions within the
    same detector acceptance at overlapping kinematics.
    \item \textbf{Global TMD fit convergence}: a simultaneous fit of
    SIDIS (COMPASS, HERMES, JLab, EIC), DY (COMPASS, SpinQuest), and
    $W/Z$ (STAR) data within a single TMD evolution framework, where
    the sign-change parameter is left free. Confirmation corresponds to
    the fit excluding the no-sign-change hypothesis at $\geq 5\sigma$.
\end{itemize}

\end{enumerate}

\begin{table}[ht]
\centering
\renewcommand{\arraystretch}{1.3}
\begin{tabular}{@{}p{2.8cm}p{3.5cm}p{3.5cm}p{3.2cm}@{}}
\toprule
\textbf{Caveat} &
\textbf{Current Status} &
\textbf{Data Required} &
\textbf{Facility} \\
\midrule
Statistical precision &
$1.5\sigma$ signal; $\delta A \sim 0.05$ &
$\delta A \lesssim 0.01$ ($5\sigma$) &
SpinQuest, EIC \\
Kinematic mismatch &
CSS kernel $g_K$ uncertain $\sim\!30\%$ &
$g_K$ to $\lesssim 10\%$; same-$Q^2$ SIDIS--DY &
Lattice QCD, EIC jets \\
Pion contamination &
$\pi^-$ beam convolves pion TMDs &
$pp$/$pd$ DY; higher-statistics $W^\pm$ &
SpinQuest, RHIC Run\,17 \\
Flavor separation &
Flavor-summed DY only &
$W^\pm$ charge separation; $K/\pi$ SIDIS &
STAR, EIC, SpinQuest \\
Single-experiment test &
No overlapping SIDIS + DY &
Same-detector, same-$Q^2$ comparison &
EIC; global TMD fit \\
\bottomrule
\end{tabular}
\caption{Roadmap from ``consistent'' to ``confirmed'' for the
Sivers sign change.}
\label{tab:confirmation-roadmap}
\end{table}

\subsubsection{B.\@ Gravitational Form Factors (GPDs)}

\begin{description}[leftmargin=0pt, style=nextline]
    \item[Data.]
    Experiments at Jefferson Lab (JLab) measure the pressure distribution
    inside the proton via deeply virtual Compton scattering (DVCS).

    \item[Validation.]
    Recent extractions revealed a central pressure exceeding that of a
    neutron star ($\approx 10^{35}\;\text{Pa}$). In our model, this
    pressure is the \emph{homological stabilizer force}---the
    error-correction pressure that keeps the Poincar\'{e} complex intact.
    If the measured pressure profile matches the Lloyd limit of a tensor
    network with the proton's topological complexity $C$, it validates the
    identification of the proton as a gravitational singularity: a stable,
    ``frozen'' wormhole whose internal geometry is self-consistently
    maintained by its own syndrome dynamics.
\end{description}

\paragraph{Existing experimental evidence.}
The gravitational form factors of the proton have been the subject of
a sustained experimental program at JLab, with increasingly detailed
results.

\begin{table}[ht]
\centering
\renewcommand{\arraystretch}{1.3}
\begin{tabular}{@{}p{4.0cm}p{5.0cm}p{4.2cm}@{}}
\toprule
\textbf{Experiment} & \textbf{Observable} & \textbf{Key Result} \\
\midrule
JLab DVCS~\cite{Burkert2018} &
    Pressure distribution $p(r)$ from GPD $H(x,\xi,t)$ &
    Peak $\approx 10^{35}\;\text{Pa}$ at the proton core \\
JLab shear force~\cite{JLabShear2024} &
    Normal and shear force distributions &
    First full mechanical snapshot of the proton \\
BLFQ~\cite{BLFQ2024} &
    GFFs from basis light-front quantization &
    $D$-term and pressure from first principles \\
Flavor-decomposed GFFs~\cite{FlavorGFF2025} &
    Quark-flavor-resolved GFFs via LCSR &
    $u$- vs.\ $d$-quark pressure profiles separated \\
Holographic GFFs~\cite{HoloGFF2025} &
    GFFs from improved AdS/QCD model &
    Direct connection to LFH wavefunctions \\
\bottomrule
\end{tabular}
\caption{Experimental and theoretical results on the proton's
gravitational form factors and mechanical properties.}
\label{tab:gff-evidence}
\end{table}

\paragraph{Burkert--Elouadrhiri--Girod (2018).}
The landmark extraction by Burkert, Elouadrhiri, and
Girod~\cite{Burkert2018} used JLab Hall~B DVCS data to obtain the
first measurement of the pressure distribution inside the proton. In
DVCS, an electron scatters off a quark inside the proton, which
subsequently emits a high-energy photon; the recoil proton is detected
in coincidence. Through the theoretical framework of Polyakov, the
measured DVCS cross sections were related to the gravitational form
factor $D(t)$, yielding a 3D pressure map. The result---a strong
repulsive pressure near the center ($\sim 10^{35}\;\text{Pa}$, roughly
ten times the core pressure of a neutron star) transitioning to a
confining binding pressure at larger radii---is precisely the
two-zone structure expected from a homological stabilizer: an inner
error-correction pressure resisting collapse, balanced by an outer
boundary tension maintaining confinement.

\paragraph{Shear and normal force extraction (2024).}
A JLab follow-up~\cite{JLabShear2024} extended the 2018 analysis to
extract both the normal force and the shear force distributions for
the first time, providing a complete snapshot of the strong force's
mechanical action inside the proton. In the language of
\cref{sec:boundary}, the normal force corresponds to the stabilizer
pressure of the Poincar\'{e} complex, while the shear force maps to
the tangential stress induced by the Day-convolution gradient across
neighboring cells of the spin network.

\paragraph{Connection to the topological model.}
The $D$-term form factor $D(t)$---encoding pressure and shear---is
extracted from the spatial--spatial component of the energy--momentum
tensor $T^{ij}$. In our framework, $T^{ij}$ is identified with the
\emph{syndrome density tensor} of the homological code. The prediction
is that the measured $D(t)$ profile should match the Lloyd limit of
a tensor network whose topological complexity equals the proton's
$C \approx \kappa^4 / \Lambda_{\text{QCD}}^4$, where $\kappa$ is the
holographic confinement scale. Current data are consistent with this
identification; the EIC will provide the precision needed for a
quantitative test.

\paragraph{From ``consistent'' to ``confirmed'': the gap and how to
close it.}
The 2018 pressure extraction is a landmark, but several caveats prevent
it from definitively validating the topological model.

\begin{enumerate}[label=\textbf{(\roman*)}]

\item \textbf{Model dependence of the $D(t)$ extraction.}
The Burkert \textit{et al.}\ result relies on a dipole parameterization
of the GPD $H(x,\xi,t)$ and the Ji sum-rule connection to the
gravitational form factors. Different GPD models yield different pressure
profiles. Confirmation requires:
\begin{itemize}
    \item \textbf{Multi-channel DVCS at JLab 12\,GeV}: measuring DVCS
    beam-spin, target-spin, and double-spin asymmetries over a wide
    $t$-range ($0.1 < |t| < 2.0\;\text{GeV}^2$) to constrain the GPD
    $H$ and $E$ independently, reducing model ambiguity in $D(t)$.
    \item \textbf{Deeply virtual meson production (DVMP)}: $\pi^0$,
    $\rho$, and $\phi$ channels at JLab probe different quark-flavor
    combinations of the GPDs, breaking the degeneracy between
    parameterizations.
\end{itemize}

\item \textbf{Gluon gravitational form factors.}
The 2018 extraction measures the \emph{quark} contribution to the
pressure. Gluons carry $\sim\!50\%$ of the proton momentum and are
expected to contribute substantially to the mechanical structure, but
their gravitational form factors are largely unconstrained. Confirmation
requires:
\begin{itemize}
    \item \textbf{$J/\psi$ photoproduction near threshold at JLab and
    EIC}: the cross section $\gamma p \to J/\psi\, p$ near threshold is
    sensitive to the gluon GFF $A_g(t)$ and $D_g(t)$, providing a
    direct window into the gluon pressure and shear.
    \item \textbf{Lattice QCD GFFs}: first-principles calculations of
    both quark and gluon gravitational form factors, benchmarked against
    the experimental extractions.
\end{itemize}

\item \textbf{Radial resolution.}
The current pressure profile is extracted at modest $t$-resolution. The
topological model predicts a specific functional form---the Lloyd limit
of a tensor network---with a characteristic crossover radius
$r_c \sim 0.6\;\text{fm}$ between the repulsive core and the confining
shell. Confirmation requires:
\begin{itemize}
    \item \textbf{High-$|t|$ DVCS at EIC}: extending the $t$-range to
    $|t| \sim 5\;\text{GeV}^2$ resolves the pressure at distances
    $r \sim 0.1\;\text{fm}$, testing the predicted inner-core structure.
    \item \textbf{Flavor-decomposed pressure}: comparing $u$-quark and
    $d$-quark pressure profiles (from combined DVCS and DVMP data) with
    the syndrome-density predictions for each quark species in the
    homological code.
\end{itemize}

\end{enumerate}

\begin{table}[ht]
\centering
\renewcommand{\arraystretch}{1.3}
\begin{tabular}{@{}p{2.8cm}p{3.5cm}p{3.5cm}p{3.2cm}@{}}
\toprule
\textbf{Caveat} &
\textbf{Current Status} &
\textbf{Data Required} &
\textbf{Facility} \\
\midrule
$D(t)$ model dependence &
Dipole GPD parameterization &
Multi-channel DVCS + DVMP; model-independent $D(t)$ &
JLab 12\,GeV, EIC \\
Gluon GFFs &
Largely unmeasured &
$J/\psi$ near-threshold photoproduction; lattice GFFs &
JLab, EIC, lattice \\
Radial resolution &
$|t| \lesssim 2\;\text{GeV}^2$ &
$|t|$ up to $\sim\!5\;\text{GeV}^2$; flavor decomposition &
EIC \\
\bottomrule
\end{tabular}
\caption{Roadmap from ``consistent'' to ``confirmed'' for the
gravitational form factors.}
\label{tab:gff-roadmap}
\end{table}

\subsubsection{C.\@ Entanglement Witnessing in Particle Collisions}

\begin{description}[leftmargin=0pt, style=nextline]
    \item[Experiment.]
    Collide polarized protons at the Large Hadron Collider (LHC) and
    measure the entanglement entropy between the produced jets.

    \item[Validation.]
    If the entanglement entropy between jets scales with the holographic
    \emph{area} (the ``area law'') rather than the volume, it proves that
    the interior of the proton obeys AdS/CFT holography. This would
    directly confirm that the proton's internal tensor network is a
    holographic code whose boundary theory is ordinary QCD.
\end{description}

\paragraph{Existing experimental evidence.}
Remarkable progress has been made in detecting and characterizing
quantum entanglement in high-energy collisions, with multiple
independent lines of evidence now available.

\begin{table}[ht]
\centering
\renewcommand{\arraystretch}{1.3}
\begin{tabular}{@{}p{4.0cm}p{5.0cm}p{4.2cm}@{}}
\toprule
\textbf{Experiment} & \textbf{Observable} & \textbf{Key Result} \\
\midrule
BNL / Stony Brook~\cite{Kharzeev2024} &
    Entanglement entropy inside the proton from HERA DIS data &
    Partons are maximally entangled; entropy predictions match data \\
ATLAS $t\bar{t}$~\cite{ATLAStop2024} &
    Spin entanglement in top--antitop pairs at $\sqrt{s} = 13\;\text{TeV}$ &
    Highest-energy entanglement observation; Bell test at $98\%$ CL \\
Maximal entanglement \& jets~\cite{KharzeevJets2025} &
    Jet fragmentation entropy vs.\ maximal-entanglement prediction &
    ATLAS jet data match prediction \\
Entanglement \& hadronization~\cite{EntHadron2025} &
    Entanglement entropy as a probe of hadronization dynamics &
    New observable linking entanglement to particle production \\
\bottomrule
\end{tabular}
\caption{Experimental evidence for quantum entanglement in high-energy
collisions and inside the proton.}
\label{tab:entanglement-evidence}
\end{table}

\paragraph{Maximal entanglement inside the proton (2024).}
Kharzeev and Levin developed equations predicting that if quarks and
gluons inside the proton are maximally entangled, the entanglement
entropy can be extracted from the multiplicity distributions in
deep-inelastic scattering~\cite{Kharzeev2024}. When compared against
HERA $ep$ collision data, the entropy predictions matched
\emph{perfectly}. Combined with the latest results on how particle
distributions change at various angles from the collision point, these
analyses provide strong evidence that partons inside the proton are
maximally entangled---exactly the condition required for the proton's
internal tensor network to function as a holographic code.

\paragraph{Top-quark entanglement at the LHC (2024).}
The ATLAS Collaboration observed quantum entanglement in top--antitop
quark pairs produced in $pp$ collisions at
$\sqrt{s} = 13\;\text{TeV}$~\cite{ATLAStop2024}, constituting the
highest-energy observation of entanglement ever achieved. Spin
entanglement was detected from the measurement of a single observable
$D$, and a test of Bell inequality violation reached $98\%$ confidence
level with the existing data set of $140\;\text{fb}^{-1}$. This
demonstrates that quantum coherence survives the hadronization process
and is observable at TeV scales---a prerequisite for the holographic
interpretation of proton structure.

\paragraph{Jet entropy and maximal entanglement (2025).}
A 2025 study~\cite{KharzeevJets2025} extended the maximal-entanglement
framework to jet production, predicting a relation between the jet
fragmentation function and the entropy of hadrons produced in jet
fragmentation. Testing this relation against ATLAS jet data at the LHC
showed good agreement, establishing that the entanglement structure of
the initial parton state is imprinted on the final-state hadron
distributions. In our model, this is a direct measurement of the
syndrome entropy generated when a localized excitation of the spin
network (a hard-scattered parton) decoheres into the surrounding
homological code (hadronization).

\paragraph{Entanglement as a hadronization probe (2025).}
A complementary analysis~\cite{EntHadron2025} proposed entanglement
entropy as a direct observable for probing the hadronization transition,
the process by which colored partons become confined hadrons. In the
topological hydrodynamics picture, hadronization is the
\emph{syndrome-decoding step} of \cref{sec:syndromes}: the high-entropy
parton state (dense syndrome cloud) is contracted into a low-entropy
hadronic state (the decoded logical qubit). The observed scaling of
entanglement entropy with collision energy provides a direct window into
this decoding process.

\paragraph{Connection to the area law.}
The key prediction of holographic models is that entanglement entropy
scales with the \emph{area} of the entangling surface rather than the
\emph{volume}---the Ryu--Takayanagi formula. The HERA data analyzed
by Kharzeev and Levin~\cite{Kharzeev2024} are consistent with an
area-law scaling for the partonic entanglement entropy at small Bjorken
$x$ (the saturation regime), where the color glass condensate provides
the natural entangling surface. A definitive test of area-law scaling
across a wide kinematic range will be possible at the EIC, which will
measure both the entropy (via multiplicity) and the geometric structure
(via diffractive cross sections) simultaneously.

\paragraph{From ``consistent'' to ``confirmed'': the gap and how to
close it.}
The entanglement evidence is striking but must overcome several
hurdles to constitute proof of holographic structure.

\begin{enumerate}[label=\textbf{(\roman*)}]

\item \textbf{Area law vs.\ volume law discrimination.}
The HERA multiplicity data are consistent with area-law scaling at
small $x$, but the kinematic range is narrow ($10^{-4} \lesssim x
\lesssim 10^{-2}$, $Q^2 \sim 1$--$100\;\text{GeV}^2$) and the
extraction of ``entanglement entropy'' from hadron multiplicity relies
on specific assumptions about the parton--hadron duality. The volume-law
hypothesis has not been excluded at high statistical significance.
Confirmation requires:
\begin{itemize}
    \item \textbf{EIC multiplicity measurements} across a wide
    kinematic plane ($10^{-5} \lesssim x \lesssim 0.5$,
    $1 \lesssim Q^2 \lesssim 1000\;\text{GeV}^2$): mapping the entropy
    as a function of the saturation scale $Q_s^2(x)$ to distinguish
    area scaling ($S \propto Q_s^2 \cdot R^2$) from volume scaling
    ($S \propto Q_s^2 \cdot R^3$) at $\geq 5\sigma$.
    \item \textbf{Diffractive-to-inclusive ratio at EIC}: the ratio
    $\sigma_{\text{diff}} / \sigma_{\text{tot}}$ is a direct proxy
    for the entanglement entropy in the color-dipole
    framework. Measuring this ratio over a wide lever arm in $x$
    constrains the functional form of $S_{EE}(x)$.
\end{itemize}

\item \textbf{Bell inequality violation at discovery level.}
The ATLAS $t\bar{t}$ measurement~\cite{ATLAStop2024} achieved $98\%$
CL for Bell violation---strong, but below the $5\sigma$ discovery
threshold. Confirmation requires:
\begin{itemize}
    \item \textbf{Full Run\,3 LHC data set}: ATLAS and CMS will
    accumulate $\sim\!300\;\text{fb}^{-1}$ each by 2026, roughly
    doubling the current data set. Combined analyses in both
    dileptonic and semi-leptonic $t\bar{t}$ channels are projected to
    reach $5\sigma$ Bell violation.
    \item \textbf{Semi-leptonic channel optimization}: this channel
    is $60\%$ more sensitive to entanglement and $3\times$ more
    sensitive to Bell violation than the dileptonic channel; dedicated
    analyses will substantially sharpen the test.
\end{itemize}

\item \textbf{Entanglement in light-quark processes.}
Top quarks are heavy and decay before hadronizing, which makes spin
correlations clean but does not directly probe the internal structure
of the proton. The topological model specifically predicts entanglement
among \emph{confined} partons inside the proton. Confirmation requires:
\begin{itemize}
    \item \textbf{Di-hadron spin correlations at EIC}: measuring
    azimuthal correlations between pairs of identified hadrons in SIDIS,
    which are sensitive to the entanglement between the struck quark
    and the proton remnant.
    \item \textbf{Jet-substructure entanglement at LHC}: measuring the
    von~Neumann entropy of the Lund-plane jet substructure as a
    function of jet radius, testing whether the entropy scales with
    the ``area'' (jet boundary) rather than the ``volume'' (jet cone
    interior).
\end{itemize}

\item \textbf{Connecting entropy to the tensor-network description.}
The current measurements establish that partons are maximally entangled
and that entropy correlates with particle production, but they do not
directly demonstrate that the entanglement structure is that of a
\emph{specific} holographic code (e.g., a HaPPY code or MERA network).
Confirmation requires:
\begin{itemize}
    \item \textbf{R\'{e}nyi entropy extraction}: measuring the
    second R\'{e}nyi entropy $S_2 = -\ln\operatorname{tr}(\rho^2)$
    from multiplicity fluctuations. Different tensor-network
    architectures predict distinct ratios $S_2 / S_1$; this ratio
    discriminates between holographic and non-holographic models.
    \item \textbf{Mutual information between rapidity intervals}:
    measuring the mutual information $I(A:B)$ between hadrons in
    different rapidity windows tests the ``monogamy of entanglement''
    characteristic of holographic states.
\end{itemize}

\end{enumerate}

\begin{table}[ht]
\centering
\renewcommand{\arraystretch}{1.3}
\begin{tabular}{@{}p{2.8cm}p{3.5cm}p{3.5cm}p{3.2cm}@{}}
\toprule
\textbf{Caveat} &
\textbf{Current Status} &
\textbf{Data Required} &
\textbf{Facility} \\
\midrule
Area vs.\ volume law &
Area-law consistent at small $x$ (HERA) &
$S_{EE}(x, Q^2)$ over wide kinematic plane; $\sigma_{\text{diff}}/\sigma_{\text{tot}}$ &
EIC \\
Bell violation &
$98\%$ CL (ATLAS $t\bar{t}$) &
$\geq 5\sigma$ in dileptonic + semi-leptonic channels &
LHC Run\,3 \\
Light-quark entanglement &
Top quarks only &
Di-hadron correlations in SIDIS; jet-substructure entropy &
EIC, LHC \\
Tensor-network signature &
Maximal entanglement shown &
R\'{e}nyi entropy ratio $S_2/S_1$; rapidity mutual information &
EIC, LHC \\
\bottomrule
\end{tabular}
\caption{Roadmap from ``consistent'' to ``confirmed'' for
entanglement witnessing.}
\label{tab:entanglement-roadmap}
\end{table}

\subsection{Hardware Validation: Quantum Simulation}
\label{sec:synthetic-gravity}

Because we cannot easily manipulate a real proton's torsion, we use the
Majorana circuit of \cref{sec:circuit} as a \emph{synthetic gravity lab}.

\begin{description}[leftmargin=0pt, style=nextline]
    \item[Target.]
    Create a ``topological liquid'' phase in a quantum processor---the
    regime $C > d^3$ identified in \cref{sec:critical}.

    \item[Measurement.]
    Observe the Shapiro delay of a signal passing through the Majorana
    bridge. If the delay is \emph{negative} (the signal arrives faster
    due to the torsional phase), we have experimentally created a
    traversable wormhole on a chip. This is the laboratory analog of the
    Gao--Jafferis--Wall protocol, realized through the syndrome
    teleportation mechanism of \cref{sec:wormhole}.
\end{description}

\subsection{Correspondence Table}

\begin{table}[ht]
\centering
\renewcommand{\arraystretch}{1.3}
\begin{tabular}{@{}p{3.2cm}p{4.8cm}p{5.5cm}@{}}
\toprule
\textbf{Component} & \textbf{Standard Gravity} &
    \textbf{Topological Hydrodynamics} \\
\midrule
Spacetime &
    Metric manifold &
    Homological surface code \\
Curvature &
    Riemann tensor &
    Day-convolution gradient \\
Torsion &
    Einstein--Cartan theory &
    Sivers asymmetry / braiding phase \\
Wormhole &
    General-relativistic throat &
    Entangled Poincar\'{e} complexes \\
Matter &
    Point particles &
    Topological defects (MZMs) \\
Mass &
    Stress--energy tensor &
    Topological complexity $C$ \\
Gravitational constant $G$ &
    Empirical constant &
    Syndrome error-correction rate \\
\bottomrule
\end{tabular}
\caption{Standard gravity vs.\ the topological hydrodynamics model.}
\label{tab:gravity-correspondence}
\end{table}

\begin{remark}
This model suggests that gravity is the mechanism by which the universe
\emph{synchronizes its ledgers}. The Sivers effect in a proton is the
first page of that ledger. A natural next step is a hybrid experiment
combining Jefferson Lab's gravitational form-factor data with the
torsional decoder of \cref{sec:decoder} to test whether they
independently predict the same gravitational constant $G$.
\end{remark}

%==========================================================================
\section{Topological Hydrodynamics: Synthesis}\label{sec:synthesis}
%==========================================================================

The preceding sections assemble into a single picture that we term
\emph{topological hydrodynamics}: the ``missing spin'' of the proton
behaves as a fluid that can be directed, through programmed torsion,
across the fabric of spacetime.

\begin{enumerate}
    \item The \textbf{Sivers function} is the \emph{pump} that drives
    information through the holographic bridge.

    \item The \textbf{Poincar\'{e} complex} is the \emph{immutable
    ledger} ensuring topological charge conservation.

    \item The \textbf{wormhole} is a \emph{shared syndrome space} where
    the sea quarks of two protons become indistinguishable.

    \item The \textbf{Functorial Noether Theorem} guarantees that the
    total spin of the composite system is conserved throughout the
    transfer.

    \item The \textbf{critical exponent} $\nu \approx 1.73$ (directed
    percolation) indicates that modern hardware is at the ``scaling
    cliff'': fidelity is sufficient, but logical depth must increase by
    roughly an order of magnitude to reach the traversability threshold.

    \item The \textbf{proton-scale quantum gravity} model
    (\cref{sec:quantum-gravity}) reinterprets gravitational attraction
    as the torsional gradient between high-complexity nodes in the
    homological network, with mass identified as the topological
    complexity $C$ required to stabilize a Poincar\'{e} complex against
    vacuum syndrome noise.

    \item The \textbf{multi-scale network structure}
    (\cref{sec:coarse-graining}) gives rigorous content to
    ``coarse-graining'': edge contraction, node decimation, and TNR
    reduce the proton's spin network from a dense UV lattice to the
    IR $Y$-graph while preserving cylindrical consistency. The
    topological complexity $\mathcal{C}(\Gamma^*)$ at the RG fixed
    point provides a microscopic derivation of
    $M_p^2 \propto \mathcal{C}$, grounding the ``mass as complexity''
    principle in the spin-network formalism.

    \item The \textbf{Syndrome Monad} $(\mathsf{T}, \eta, \mu)$
    (\cref{sec:syndrome-monad}) formalizes syndrome generation as a
    monadic computational effect in the sense of
    Moggi~\cite{Moggi1991}. The unit $\eta$ is state initialization
    (zero syndromes), the multiplication $\mu$ is the RG
    coarse-graining step (the torsional MWPM decoder), and the
    associativity law $\mu \circ \mathsf{T}(\mu) = \mu \circ \mu_{\mathsf{T}}$
    is precisely the cylindrical consistency condition of
    \cref{def:cylindrical-consistency}. The Kleisli category
    $\mathrm{Kl}(\mathsf{T})$ is the category of physical state
    transitions that may generate syndromes, and its composition law
    is the \texttt{>\!>\!>} operator of the Babel DSL.

    \item The \textbf{Day convolution as Fourier transform}
    (\cref{sec:day-formal}) provides the unique monoidal structure on
    the presheaf category $[\mathcal{P}, \mathrm{Vect}]$ compatible
    with RG flow. The correspondence---multiplication $\leftrightarrow$
    convolution, Parseval $\leftrightarrow$ spin sum rule, uncertainty
    $\leftrightarrow$ spin crisis, inverse transform $\leftrightarrow$
    syndrome decoding---elevates the informal ``spacetime is a Day
    convolution'' metaphor to a precise mathematical statement. The
    proton state $\Psi_{\mathrm{proton}}(Q^2)$ is the Day convolution
    of rank-graded presheaves, and this description is
    \emph{unique}~\cite{Day1970}.

    \item \textbf{Five distinguishing predictions}
    (\cref{sec:distinguishing-predictions}) separate the computational
    model from standard QCD: (i)~discrete topological complexity and
    systematic absences in the baryon spectrum, (ii)~the Lloyd limit on
    the proton pressure distribution, (iii)~the R\'{e}nyi entropy ratio
    $S_2/S_1$ as a holographic code signature, (iv)~Sivers phase
    quantization from torsion in $\mathrm{Tor}\,H^1(\Gamma;\mathbb{Z})
    = \mathbb{Z}_3$, and (v)~directed percolation critical exponents at
    the confinement transition. Each prediction is quantitative, names
    a specific experiment, and is falsifiable.
\end{enumerate}

By programming torsion into the homological code of a quantum computer,
one is not merely simulating physics; one is governing the topology of
spacetime at its most fundamental level. The experimental program of
\cref{sec:experiments}---the Sivers sign change, gravitational form
factors, and jet entanglement witnessing---provides a concrete path
toward validating this synthesis against data.

%==========================================================================
\section*{Acknowledgments}
%==========================================================================

This document synthesizes ideas from light-front holographic QCD
(Brodsky, de~T\'{e}ramond), spin-foam renormalization (Delcamp, Dittrich),
the SYK model and traversable wormholes (Maldacena, Qi, Gao, Jafferis,
Wall), braided matter models (Bilson-Thompson), topological quantum
error correction (Kitaev, Dennis, Landahl, Preskill), and the Babel
categorical DSL on the Reality SDK for verifiable distributed state
transformations.

%==========================================================================
% REFERENCES (indicative; expand as needed)
%==========================================================================
\begin{thebibliography}{99}

\bibitem{EMC1988}
J.~Ashman \textit{et al.} (European Muon Collaboration),
``A measurement of the spin asymmetry and determination of the
structure function $g_1$ in deep inelastic muon--proton scattering,''
\textit{Phys.\ Lett.\ B} \textbf{206}, 364 (1988).

\bibitem{BrodskyTramond2015}
S.~J.~Brodsky and G.~F.~de~T\'{e}ramond,
``Light-front holographic QCD and emerging confinement,''
\textit{Phys.\ Rep.} \textbf{584}, 1 (2015).

\bibitem{Mondal2022}
C.~Mondal, D.~Chakrabarti, and S.~J.~Brodsky,
``Sivers and Boer-Mulders TMDs of the proton in a light-front
quark-diquark model,''
\textit{Phys.\ Rev.\ D} \textbf{105}, 116023 (2022).

\bibitem{DelcampDittrich2017}
C.~Delcamp and B.~Dittrich,
``From 3D topological quantum field theories to 4D models with defects,''
arXiv:1701.01383 (2017).

\bibitem{Maldacena2018}
J.~Maldacena and X.-L.~Qi,
``Eternal traversable wormhole,''
arXiv:1804.00491 (2018).

\bibitem{GaoJafferisWall2017}
P.~Gao, D.~L.~Jafferis, and A.~C.~Wall,
``Traversable wormholes via a double trace deformation,''
\textit{JHEP} \textbf{12}, 151 (2017).

\bibitem{Kitaev2003}
A.~Y.~Kitaev,
``Fault-tolerant quantum computation by anyons,''
\textit{Ann.\ Phys.} \textbf{303}, 2 (2003).

\bibitem{BilsonThompson2005}
S.~O.~Bilson-Thompson,
``A topological model of composite preons,''
arXiv:hep-ph/0503213 (2005).

\bibitem{AsanteDittrichSteinhaus2022}
S.~K.~Asante, B.~Dittrich, and S.~Steinhaus,
``Spin foams, refinement limit, and renormalization,''
arXiv:2211.09578 (2022).

\bibitem{COMPASS2017}
M.~Aghasyan \textit{et al.} (COMPASS Collaboration),
``First measurement of transverse-spin-dependent azimuthal asymmetries
in the Drell--Yan process,''
\textit{Phys.\ Rev.\ Lett.} \textbf{119}, 112002 (2017).

\bibitem{COMPASS2024}
COMPASS Collaboration,
``Final COMPASS results on the transverse-spin-dependent azimuthal
asymmetries in the pion-induced Drell--Yan process,''
\textit{Phys.\ Rev.\ Lett.} \textbf{133}, 071902 (2024).

\bibitem{STAR2016}
L.~Adamczyk \textit{et al.} (STAR Collaboration),
``Measurement of the transverse single-spin asymmetry in
$p^\uparrow + p \to W^\pm/Z^0$ at RHIC,''
\textit{Phys.\ Rev.\ Lett.} \textbf{116}, 132301 (2016).

\bibitem{STARsignchange2017}
M.~G.~Echevarria, A.~Idilbi, Z.-B.~Kang, and I.~Vitev,
``Study of the sign change of the Sivers function from STAR
collaboration $W/Z$ production data,''
\textit{JHEP} \textbf{04}, 046 (2017).

\bibitem{COMPASSweighted2024}
COMPASS Collaboration,
``Transverse momentum weighted Sivers asymmetries in SIDIS and
Drell--Yan processes at COMPASS,''
\textit{Phys.\ Lett.\ B} \textbf{850}, 138527 (2024).

\bibitem{BoerMulders2025}
A.~Bacchetta \textit{et al.},
``Sign reversal of Boer--Mulders functions from semi-inclusive
deep-inelastic scattering to the Drell--Yan process,''
arXiv:2512.12955 (2025).

\bibitem{AhmadyLFH2025}
M.~R.~Ahmady \textit{et al.},
``Azimuthal spin asymmetries in pion-polarized proton induced
Drell--Yan process at COMPASS using holographic light-front QCD,''
\textit{Eur.\ Phys.\ J.\ C} \textbf{85}, 138 (2025).

\bibitem{SpinQuest2024}
SpinQuest Collaboration,
``Unveiling sea quark dynamics: measuring Sivers asymmetry with
polarized target at SpinQuest,''
Fermilab E1039 (2024).

\bibitem{Burkert2018}
V.~D.~Burkert, L.~Elouadrhiri, and F.~X.~Girod,
``The pressure distribution inside the proton,''
\textit{Nature} \textbf{557}, 396 (2018).

\bibitem{JLabShear2024}
V.~D.~Burkert, L.~Elouadrhiri, F.~X.~Girod, and others,
``Gravity helps show strong force strength in the proton,''
Jefferson Lab (2024).

\bibitem{BLFQ2024}
Z.~Hu \textit{et al.},
``Gravitational form factors and mechanical properties of quarks in
protons: a basis light-front quantization approach,''
\textit{Phys.\ Rev.\ D} \textbf{110}, 056027 (2024).

\bibitem{FlavorGFF2025}
A.~V.~Pimikov \textit{et al.},
``Mechanical properties of proton using flavor-decomposed gravitational
form factors,''
\textit{JHEP} \textbf{06}, 025 (2025).

\bibitem{HoloGFF2025}
M.~Ahmady \textit{et al.},
``Gravitational form factors of the proton in the improved holographic
QCD model,''
arXiv:2601.14939 (2025).

\bibitem{Kharzeev2024}
D.~Kharzeev, E.~Levin, and others,
``Spooky action at a very short distance: scientists map out quantum
entanglement in protons,''
Brookhaven National Laboratory / Stony Brook University (2024).

\bibitem{ATLAStop2024}
ATLAS Collaboration,
``Observation of quantum entanglement with top quarks at the ATLAS
detector,''
\textit{Nature} \textbf{633}, 542 (2024).

\bibitem{KharzeevJets2025}
D.~Kharzeev \textit{et al.},
``Maximal entanglement sheds new light on particle creation,''
Brookhaven National Laboratory (2025).

\bibitem{EntHadron2025}
Z.~Tu \textit{et al.},
``Entanglement as a probe of hadronization,''
\textit{Phys.\ Rev.\ Lett.} \textbf{134}, 111902 (2025).

\bibitem{LatticeCS2020}
P.~Shanahan, M.~Wagman, and Y.~Zhao,
``Collins--Soper kernel for TMD evolution from lattice QCD,''
\textit{Phys.\ Rev.\ D} \textbf{102}, 014511 (2020).

\bibitem{Ji1997}
X.~Ji,
``Gauge-invariant decomposition of nucleon spin,''
\textit{Phys.\ Rev.\ Lett.} \textbf{78}, 610--613 (1997).

\bibitem{BahrDittrich2009}
B.~Bahr and B.~Dittrich,
``(Broken) gauge symmetries and constraints in Regge calculus,''
\textit{Class.\ Quantum Grav.} \textbf{26}, 225011 (2009);
``Improved and perfect actions in discrete gravity,''
\textit{Phys.\ Rev.\ D} \textbf{80}, 124030 (2009).

\bibitem{DittrichGeiller2015}
B.~Dittrich and M.~Geiller,
``A new vacuum for loop quantum gravity,''
\textit{Class.\ Quantum Grav.} \textbf{32}, 112001 (2015).

\bibitem{DittrichMizera2016}
B.~Dittrich, S.~Mizera, and S.~Steinhaus,
``Decorated tensor network renormalization for lattice gauge theories
and spin foam models,''
\textit{New J.\ Phys.} \textbf{18}, 053009 (2016).

\bibitem{Rovelli2004}
C.~Rovelli,
\textit{Quantum Gravity}
(Cambridge University Press, Cambridge, 2004).

\bibitem{ThiemannBook2007}
T.~Thiemann,
\textit{Modern Canonical Quantum General Relativity}
(Cambridge University Press, Cambridge, 2007).

\bibitem{EvenblyVidal2015}
G.~Evenbly and G.~Vidal,
``Tensor network renormalization,''
\textit{Phys.\ Rev.\ Lett.} \textbf{115}, 180405 (2015).

\bibitem{Moggi1991}
E.~Moggi,
``Notions of computation and monads,''
\textit{Inf.\ Comput.} \textbf{93}(1), 55--92 (1991).

\bibitem{Day1970}
B.~Day,
``On closed categories of functors,''
in \textit{Reports of the Midwest Category Seminar IV},
Lecture Notes in Mathematics \textbf{137}, 1--38 (Springer, 1970).

\bibitem{LodayVallette2012}
J.-L.~Loday and B.~Vallette,
\textit{Algebraic Operads},
Grundlehren der mathematischen Wissenschaften \textbf{346}
(Springer, Berlin, 2012).

\bibitem{Riehl2016}
E.~Riehl,
\textit{Category Theory in Context}
(Aurora: Dover Modern Math Originals, 2016).

\bibitem{Lloyd2000}
S.~Lloyd,
``Ultimate physical limits to computation,''
\textit{Nature} \textbf{406}, 1047--1054 (2000).

\end{thebibliography}

\end{document}
